### 2018-03-20 MS-new/SignedDepth_08_07_16_1.tex

\documentclass[12pt,letterpaper]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{authblk}
\usepackage{bibentry}
\usepackage{color}
\usepackage{epsfig}
\usepackage{fullpage}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{lmodern}
\usepackage{pdflscape}
\usepackage{pdfpages}
\usepackage{relsize}
\usepackage{scalerel}
\usepackage{setspace}
\usepackage{soul}
\usepackage{slantsc}
\usepackage{stackengine}
\usepackage{subfigure}
\usepackage{subfiles}
% \usepackage{tikz}
\usepackage{verbatim}
% \usepackage{amsmath,amssymb,amsthm}
% \usepackage{rotating}
\usepackage{wrapfig}
%\usepackage{bibentry}

\usepackage{natbib}

\newcommand{\bfa}{{\mathbf {a}}}
\newcommand{\bfb}{{\mathbf {b}}}
\newcommand{\bfc}{{\mathbf {c}}}
\newcommand{\bfd}{{\mathbf {d}}}
\newcommand{\bfe}{{\mathbf {e}}}
\newcommand{\bff}{{\mathbf {f}}}
\newcommand{\bfg}{{\mathbf {g}}}
\newcommand{\bfh}{{\mathbf {h}}}
\newcommand{\bfi}{{\mathbf {i}}}
\newcommand{\bfj}{{\mathbf {j}}}
\newcommand{\bfm}{{\mathbf {m}}}
\newcommand{\bfn}{{\mathbf {n}}}
\newcommand{\bfp}{{\mathbf {p}}}
\newcommand{\bfr}{{\mathbf {r}}}
\newcommand{\bfs}{{\mathbf {s}}}
\newcommand{\bft}{{\mathbf {t}}}
\newcommand{\bfu}{{\mathbf {u}}}
\newcommand{\bfv}{{\mathbf {v}}}
\newcommand{\bfw}{{\mathbf {w}}}
\newcommand{\bfx}{{\mathbf {x}}}
\newcommand{\bfy}{{\mathbf {y}}}
\newcommand{\bfz}{{\mathbf {z}}}

\newcommand{\bfA}{{\mathbf {A}}}
\newcommand{\bfB}{{\mathbf {B}}}
\newcommand{\bfC}{{\mathbf {C}}}
\newcommand{\bfD}{{\mathbf {D}}}
\newcommand{\bfE}{{\mathbf {E}}}
\newcommand{\bfF}{{\mathbf {F}}}
\newcommand{\bfG}{{\mathbf {G}}}
\newcommand{\bfH}{{\mathbf {H}}}
\newcommand{\bfI}{{\mathbf {I}}}
\newcommand{\bfJ}{{\mathbf {J}}}
\newcommand{\bfK}{{\mathbf {K}}}
\newcommand{\bfL}{{\mathbf {L}}}
\newcommand{\bfM}{{\mathbf {M}}}
\newcommand{\bfP}{{\mathbf {P}}}
\newcommand{\bfQ}{{\mathbf {Q}}}
\newcommand{\bfR}{{\mathbf {R}}}
\newcommand{\bfS}{{\mathbf {S}}}
\newcommand{\bfT}{{\mathbf {T}}}
\newcommand{\bfU}{{\mathbf {U}}}
\newcommand{\bfV}{{\mathbf {V}}}
\newcommand{\bfW}{{\mathbf {W}}}
\newcommand{\bfY}{{\mathbf {Y}}}
\newcommand{\bfX}{{\mathbf {X}}}
\newcommand{\bfZ}{{\mathbf {Z}}}

\newcommand{\bfalpha}{{\boldsymbol{\alpha}}}
\newcommand{\bfbeta}{{\boldsymbol{\beta}}}
\newcommand{\bfgamma}{{\boldsymbol{\gamma}}}
\newcommand{\bfdelta}{{\boldsymbol{\delta}}}
\newcommand{\bfepsilon}{{\boldsymbol{\epsilon}}}
\newcommand{\bfeta}{{\boldsymbol{\eta}}}
\newcommand{\bflambda}{{\boldsymbol{\lambda}}}
\newcommand{\bfmu}{{\boldsymbol{\mu}}}
\newcommand{\bfnu}{{\boldsymbol{\nu}}}
\newcommand{\bfomega}{{\boldsymbol{\omega}}}
\newcommand{\bfpi}{{\boldsymbol{\pi}}}
\newcommand{\bfpsi}{{\boldsymbol{\psi}}}
\newcommand{\bfsigma}{{\boldsymbol{\sigma}}}
\newcommand{\bftheta}{{\boldsymbol{\theta}}}
\newcommand{\bfxi}{{\boldsymbol{\xi}}}

\newcommand{\bfLambda}{{\boldsymbol{\Lambda}}}
\newcommand{\bfPi}{{\boldsymbol{\Pi}}}
\newcommand{\bfSigma}{{\boldsymbol{\Sigma}}}
\newcommand{\bfTau}{{\boldsymbol{T}}}
\newcommand{\bfTheta}{{\boldsymbol{\Theta}}}

\newcommand{\BE}{\mathbb{E}}
\newcommand{\BF}{\mathbb{F}}
\newcommand{\BG}{\mathbb{G}}
\newcommand{\BI}{\mathbb{I}}
\newcommand{\BR}{\mathbb{R}}
\newcommand{\BV}{\mathbb{V}}

\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cX}{\mathcal{X}}

\def\baq#1\eaq{\begin{align}#1\end{align}}

\def\ban#1\ean{\begin{align*}#1\end{align*}}

\def\bredbf#1\eredbf{{\color{red}{\bf ???? #1 ????}}}

\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Proposition}{Proposition}[section]
\newtheorem{Corollary}{Corollary}[section]
\newtheorem{Definition}{Definition}[section]
\newtheorem{Algorithm}{Algorithm}[section]
\newtheorem{Claim}{Claim}[section]

%\voffset=0.5in

   \newtheoremstyle{Example}{\topsep}{\topsep}%
     {}%         Body font
     {}%         Indent amount (empty = no indent, \parindent = para indent)
     {\bfseries}% Thm head font
     {:}%        Punctuation after thm head
    {0.9mm}%     Space after thm head (\newline = linebreak)
     {\thmname{#1}\thmnumber{ #2}\thmnote{(\it #3)}}%         Thm head spec

%	\AtEndEnvironment{Example}{\null\hfill\qed}%
   \theoremstyle{Example}
   \newtheorem{Example}{Example}[section]
	\AtEndEnvironment{Example}{\null\hfill\qed}%

\renewcommand{\baselinestretch}{1.1}

%
%\pagestyle{myheadings}
%\markboth{}{\underline{{\bf Notes: (do not circulate)} \hspace{4.5cm} {\sc  Ansu Chatterjee} \hspace{0.25cm}}}

\begin{document}
\bibliographystyle{plain}

\makeatletter
\def\@oddfoot{[\jobname (\today)]\hfil\thepage}
\makeatother

\title{On Combinations of Sign and Peripherality Transformations for Robust Analysis
of Hilbert-valued data}

%\title{Robust methods  based on the kernel signed peripherality 
%transformations}
\author{SM, SC}

\maketitle

\section{Introduction}

\bredbf
I will write this section once the rest of the work is complete
\eredbf

The scope of Hilbert-valued data is considerable, and encompasses 
the univariate and multivariate finite-dimensional observations traditionally 
studied in Statistics, as well  as certain kinds of functional data. For 
$\BR^{p}$-valued data, the classical approach is to model the joint distribution 
of the observations using a probability distribution known up to finite dimensional
parameters. If the correct probability distribution is used, there are often 
theoretical optimality guarantees. However in practice, the correct distribution is 
rarely, if ever, known; there may be aberrant observations that do not follow the 
pattern of the bulk of the data but whose influence on parameter estimates and 
eventual inference can be considerable; computational practicability and costs 
and complexity of patterns in the data may limit choices of statistical methods 
to adopt. The analysis of functional data is in a developmental stage, 
but the above comments on practical issues are applicable for such data also.

One way of circumventing the practical issues outlined above is to use 
{\it robust} techniques of data analysis, where theoretical optimality under 
restrictive idealized conditions is not of interest. Instead, efforts are made to 
make the resulting estimation and inference insensitive to  technical assumptions, 
outliers in the data, and computationally feasible. Naturally, subject to these 
robustness guarantees, the goal is to use the most efficient techniques possible.
The following very simple example recapitulates some very well known facts, but 
serves to illustrate the main drive of this paper.

\begin{Example}[Testing for a location parameter]
\label{Example:TestingLocation1d}

Suppose we have real valued observations $Y_{1}, \ldots, Y_{n}$, which from the 
domain of study that gave rise to these data, are known to be distributed around 
an unknown parameter $\mu \in \BR$, and the scientific question of interest lies in 
evaluating  whether the hypothesis $\mu = \mu_{0}$ for some given constant 
$\mu_{0}$ is tenable or not.

Under the assumption that the observations $Y_{1}, \ldots, Y_{n}$ are 
independent, identically distributed (i.i.d. hereafter) as $N (\mu, \sigma^{2})$, the 
Gaussian distribution with mean $\mu$ and variance $\sigma^{2}$, the optimal 
procedure is to use the $t$-test based on the statistic 
\ban 
T_{n} = n^{1/2} (\bar{Y}_{n} - \mu_{0})/S_{n-1}
\ean
and the Student's $t$-distribution. Here 
$\bar{Y}_{n} = n^{-1} \sum_{i=1}^{n} Y_{i}$ and 
$S_{n- 1}^{2} = (n -1)^{-1} \sum_{i=1}^{n} (Y_{i} - \bar{Y}_{n})^{2}$. 
However, note that the assumptions of the data being  independent, or identically 
distributed, or following the Gaussian distribution, or having the same variance for 
all observations, are rarely justified from the domain of study, and are merely choices of 
convenience. Also, one or more outlying observations among 
$Y_{1}, \ldots, Y_{n}$ can have extreme effect on either or both of 
$\bar{Y}_{n}$ and $S_{n- 1}^{2}$, and result in a numeric value that  completely 
misrepresents the nature of the data and misleading inference relating to the scientific query.

Alternatives to the above is to use the {\it sign}, and possibly also the 
{\it rank} of each observation. Let $\cI_{A}$ be the indicator of  $A$, 
that is, $\cI_{A} = 1$ if $A$ is true and is zero otherwise.
Define the following:
\ban 
(a) \text{ Sign function } & S (y; \nu) 
= | y - \nu |^{-1} (y - \nu) \cI_{\{ y \ne \nu \}}, \\
(b) \text{ Rank function } & R (y; Y_{1}, \ldots, Y_{n}) 
= \sum_{i=1}^{n} \cI_{ \{ y \leq Y_{i} \}}.
\ean
Then $S_{i} = S (Y_{i}; \mu_{0})$ indicates whether the $i^{th}$ observation is 
greater than $\mu_{0}$ or not, and $R_{i} = R (Y_{i}; Y_{1}, \ldots, Y_{n})$ is 
the rank that the $i^{th}$ observation takes among the entire data when sorted 
in an ascending order.

The {\it sign test statistic} corresponding to the above location testing problem
is given by $S = \sum_{i = 1}^{n} S_{i}$ and under the very general assumption 
that the $Y_{i}$'s are independent with distribution function $F_{i}$, and all the 
$F_{i}$'s have a common median $\mu_{0}$, the null distribution of 
$S$ is Binomial $(n, 0.5)$. This testing procedure is also indeed optimal in a 
very broad framework, see 
\bredbf
Need the appropriate reference from hajek, also need the conditions from there 
\eredbf

However, in many situations some additional information is known beyond the 
extremely relaxed conditions required by the sign test, and the Wilcoxon 
signed-rank test,  is found to be considerably more efficient than the sign test. 
These are well known results from several decades back, see 
\bredbf
Standard Hollander and Wolfe type refernce
\eredbf
for details.

\end{Example}

Example~\ref{Example:TestingLocation1d} motivates the present paper: which is a 
study on combinations of the equivalents of the sign function and rank function 
in Hilbert spaces, that leads to robust statistical procedures of considerable efficiency 
in a wide variety of problems. Very interestingly, we find that for location 
parameters, using in conjunction with the sign function a weight function that 
assigns {\it more} mass to observations closer to the center, results in   robust  
and considerably efficient estimators and test statistics. On the other hand, for 
scale or dispersion parameters, in conjunction with the sign function using a weight 
function that assigns {\it less} mass to observations closer to the center, 
results in  similar properties.

Notice that the robust and potentially optimal procedures outlined in 
Example~\ref{Example:TestingLocation1d}  above are based on transformations of the 
original data that preserved the structures of interest. In the rest of this paper we 
study equivalent transformations on general Hilbert spaces, and use the transformed data 
for ($i$) robust location estimation, ($ii$) testing for a location parameter in 
high dimensions and Hilbert spaces, ($iii$) robust dispersion estimation, ($iv$) 
robust principal component analysis (RPCA), ($v$) RPCA in dimension reduction and 
learning, ($vi$) robust analysis of functional data.

Given the considerable breadth of these topics, we present algorithmic details, some 
illustrative examples, and a selection of theoretical results. Considerably deeper studies 
in any of the above topics is feasible, however we postpone deep-dives into any 
specific application to future papers in order to make the main methodological drivers 
of this paper more easily understandable.

In Section~\ref{Section:SP} we present the {\it sign function} and the 
{\it peripherality function} for Hilbert spaces. The former is well known, and variants of 
the latter are also well known  in the literature. One major contribution of this 
paper is in the proposal of combining the two in different ways to achieve 
simultaneous robustness and efficiency targets. Then on the following sections we 
present details on the six applications listed above.

\section{The sign and peripherality functions}
\label{Section:SP}

We consider data $X_{1}, \ldots, X_{n}$ from some subset $\cX$ of 
a real separable Hilbert $\cH$.

We consider two functions defined below. First, we consider the 
{\it sign function} $S : \cH \times \cH \rightarrow \cH$, defined as 
\ban
S (x; \mu_{x}) = || x - \mu_{x} ||^{-1} (x - \mu_{x}) {\cI}_{ \{ x \ne \mu_{x} \}}. 
\ean
This sign function is defined with respect to the {\it location parameter} 
$\mu_{x} \in \cH$, and the norm $|| \cdot ||$ used above is the norm of the 
underlying Hilbert space. This is a direct  generalization of the 
real-valued case of the indicator of whether the point $x$ is to the right, 
left or at $\mu_{x}$, described in Example~\ref{Example:TestingLocation1d}. 
This function has been used many times in statistics, see
\bredbf insert several references. \eredbf

We next describe the {\it peripherality function}, for which some mathematical 
preliminaries are necessary for easier exposition. Let $(\Omega, \cA, \alpha)$ 
be a probability space, and let $\cB$ be the Borel $\sigma$-algebra generated 
by the norm topology of $\cH$. A $\cH$-valued random variable is a mapping 
$X :  \Omega \rightarrow \cH$ such that for every $B \in \cB$, $X^{-1} (B) \in \cA$. 
It is easy to see that $\alpha_{x} = \alpha ( X^{-1} (\cdot ) )$ is a probability 
measure on the measurable space $(\cH, \cB)$. Mathematical details about 
such probability measures on Hilbert spaces are available from a number of 
places, including
\bredbf
BLSP notes (my primary reference), Gross, Segal, and what not.
\eredbf

Let $\cM$ be a set of probability measures on $\cH$. 
%When $\cH$ 
%is infinite-dimensional, we further impose the conditions that for any $\nu \in \cM$
%\ban 
%\int_{\cH} || x || \nu (d x) < \infty, \hspace{1 cm} 
%\int_{\cH} || x ||^{2} \nu (d x) < \infty.
%\ean
A {\it peripherality function} 
$ P : \cH \times \cM \rightarrow \BR$,  is a function that satisfies the following 
condition:\\
{\it For every probability measure $F \in \cM$, there exists a constant $\mu_{F} \in \cH$ 
such that for every $t \in [ 0, 1]$ and every $x \in \cH$}
\ban 
P \Bigl( \mu_{F} ; F \Bigr) \leq P \Bigl( \mu_{F} + t ( x - \mu_{F} ); F \Bigr). 
\ean

That is, for every fixed $F$, the peripherality function achieves a minimum at 
$\mu_{F}$, and is non-decreasing in every direction away from $\mu_{F}$. If 
we impose the practical restriction that $\inf_{x} P ( x ; F )$ is finite and 
bounded below, then we may as well impose without loss of generality 
$P ( \mu_{F} ; F ) = 0$ and consequently $P ( x ; F ) \geq 0$ for all $x \in \cH$ 
and  $F \in \cM$. In many cases of interest, $P ( \cdot; \cdot)$ is 
uniformly bounded above as well.

The peripherality function quantifies whether the point $x$ is near or far from $\mu_{F}$. 
We will impose additional conditions on this function 
as we proceed, but it can be seem immediately that any distance measure between 
$x$ and $\mu_{F}$ satisfies the bare minimum requirement mentioned above.

In this paper, we demonstrate certain interesting applications arising from 
composing the sign function and the peripherality function together, to form the 
{\it signed-peripherality function} $\kappa (\cdot)$. We define this function 
with three parameters $\mu_{x} \in \cH$, $F \in \cM$ and  
$\mu_{y} \in \cH$,  argument $x \in \cH$ and range $\cH$. More precisely,
we use two functions $\kappa_{s} : \cH \rightarrow \cH$, 
$\kappa_{p} : \cH \rightarrow \cH$ that are respectively 
composed with the sign transformation and the peripherality function, and 
then multiplied together to obtain the function
$\kappa : \cH \times \cH \times \cM \times \cH \times \cH \rightarrow \cH$ 
defined as 
\ban 
\kappa (x; \mu_{x}, F, \mu_{y} ) = k_{s} (S (x; \mu_{x})) k_{p} (P (x; F)) + \mu_{y}. 
\ean
Within this very generic framework, we will explore two simple choices. We consider 
$\kappa_{s} (x) = x$, thus this is fixed to be the identity transformation. The two 
alternatives we consider for $\kappa_{p}$ are 
$\kappa_{p} (x) = x$ and $\kappa_{p} (x) = \exp (- x)$, thus one is linearly 
increasing with $x$ while the other is exponentially decreasing.

Notice that if we consider $\mu_{y} = \mu_{F} = \mu_{x}$,  
$\kappa_{s} (x) = \kappa_{p} (x) = x$, and take the very 
simple peripherality  function $P( x; F) = || x - \mu_{F} ||$, we have 
$\kappa (x; \mu_{x}, F, \mu_{y} ) \equiv x$ for all choices of parameters 
$\mu_{x}, F, \mu_{y}$.  Consequently, under this choice of parameters for the 
$\kappa$-transformation, analyzing a dataset $\{ X_{1}, \ldots, X_{n} \}$ and 
its $\kappa$-transformed version 
$\{ Y_{i} = \kappa (X_{i}; \ldots), \ i = 1, \ldots, n \}$ are equivalent. However,
in this paper we illustrate how other choices of the peripherality function 
lead to interesting robustness results. We have deliberately set the location 
parameters $\mu_{x}, \mu_{F}, \mu_{y}$ to be potentially non-identical, this 
additional flexibility has some advantage for robust data analysis. In many 
applications, the value of these three parameters may be identical, which leads 
to no conflict in our framework.

A whole class of peripherality functions can be defined from 
 {\it data - depth},  which are center-outward ranking of multivariate data. 
 Data-depths have been extensively used in statistics also, see 
 \bredbf multiple references. \eredbf
 Peripherality functions can be defined as some inverse ranking based 
 on data depth, and the concept of {\it outlyingness} associated with data depth 
 is essentially same as what we use in this paper. We use the term 
 {\it peripherality} to keep track of the difference in application 
 contexts and technical assumptions. As an additional objective of this paper, 
 we discuss  some properties and uses of  data-depth in real, separable Hilbert spaces.

 {\bf old material:}{\it
 In this paper, we consider a few illustrative cases of the use of the 
 $\kappa$-transformation. Suppose the data at hand is $X_{1}, \ldots, X_{n}$, 
 and we define $Y_{i} = \kappa (X_{i}; \mu_{X}, F, \mu_{Y})$ for some 
choice of parameters $\mu_{X}, F, \mu_{Y}$. For interpretability and 
convenience, we assume that $\BE S ( X_{i}; \mu_{X}) P ( X_{i}; F) = 0$, thus 
$\BE Y_{i} = \mu_{Y}$. We thus have 
\ban 
\BV Y_{i} 
& = \BE P ( X_{i}; F)^{2} S ( X_{i}; \mu_{X}) S ( X_{i}; \mu_{X})^{T} \\
& = \BE P ( X_{i}; F)^{2} || X_{i} - \mu_{X} ||^{-2} 
( X_{i} - \mu_{X} ) (X_{i} - \mu_{X})^{T}.
\ean
}

\bredbf Need to include (a) Biman-PC idea for affine equivariance for $p \ll n$, 
(b) kernel versions as an example of generalization. (c) anything else? \eredbf

\section{The robust location problem}

%\bredbf
%This section and the section following this needs to be merged and made sensible and 
%publication-worthy. 
%\eredbf

\subsection{General approach}

Here are things we need to do:
\begin{enumerate}

\item Propose robust location estimation in  a general Hilbert space $\cH$, based on 
$\kappa_{s} (x) = x$ and $\kappa_{p} (x) = \exp (-x)$. I am using $\exp (-x)$ and not 
$1/x$ (on which you have done some simulations successfully I think) so that we do 
not have unboundedness issues.

\item Suggest that this may be a better location estimate than the spatial median 
that is obtained by just minimizing the sign function.

\item Show whatever theoretical results we have. Make these completely error-free, 
so no handwaving or loose, non-crisp arguments.

\item Do the infinite dimensional case also. Our strongest point in this paper is being 
able to handle robustness in infinite dimensions. There may not be any useful 
or usable definition of invariance or equivariance in infinite dimensions. So separate 
the results for the $p \ll n$ case and the high-dimensional case, if necessary.

\item Show simulation results.

\item Do one running data analysis in infinite dimensions. Working with extreme 
precipitation or temperatures is a good idea, there is a potentially a need for robust  
location estimation there.

\item Let us finish one section first, then move to another section.
\end{enumerate}

\subsection{The following are concrete suggestions:}

\begin{enumerate}

\item 
%\bredbf 
Describe what is being observed, ie, the main aspects of the data. For example, are 
you assuming iid data throughout?
%\eredbf

\item 
%\bredbf
After you have described the data, define the parameter you are interested in. 
The population mean, population median, 
the population equivalent of the minimizer of the weighted sign function are not all 
same. However, each of them can be a candidate for a {\it location parameter}. Don't skip 
the mathematical details, include everything.
%\eredbf

\item 
%\bredbf 
Then propose your estimator for the parameter of your choice. 
%\eredbf

\item 
%\bredbf
Now explore properties of your proposed estimator. Simulate data and show that 
your estimator is better than the (sample) mean in terms of robustness, and 
better than the (sample) median in terms of efficiency. The following steps 
describe what you might simulate.
%Use curves as data, so 
%we are justified in using the term ``Hilbert-valued''. 
%\eredbf

\item Use the el Nino data as a template. Do your robust PCA, and extract the first 
5 or so PCs with their loadings $\beta_{1}, \beta_{2}, \ldots, \beta_{5}$. This will form
the foundation for the simulation. For the time index of the data, make it $t \in [0, 1)$
to avoid technical difficulties. That is, consider June as $t = 0$, July as $t = 1/12$ and 
so on. This simulation will also form the spine of the paper for each application (we will use this simulated data for all the sections), so do this carefully. Use {\tt set.seed}, so we can reproduce results.

\item
We will generate $n$ curves indexed by $i = 1, \ldots, n$, 
all of which will be sampled at time points 
%$t_{i k}$, $k =1, \ldots, T_{i}$. Scale the total 
%time to be between $[0, 1]$, so that there are no technical difficulties. You may take 
%$T_{1} = \ldots = T_{n} = T$ for convenience, and in fact even take 
$t_{1}, \ldots, t_{T}$. That is, the observation times for all the curves are 
same. You can take $T = 12$ and $t_{j} = (j -1)/12$.

\item 
For simulation, first generate $b_{i j} \sim N ( \beta_{j}, V_{j})$ for $j = 1, \ldots, 5$, 
and define the simulated signal for the $i^{th}$ curve as  
\ban 
\xi_{i}  (t) = \sum_{j = 1}^{5} b_{i j} PC_{j} (t), \ i=1, \ldots, n. 
\ean
Use $V_{2}$ very large, so that in different signals there is considerable variability 
in the signal.

\item
Add a iid high-variance noise, say $t_{2}$ or something to each observed data point 
$t_{i k}$, $k =1, \ldots, T_{i}$, $i =1, \ldots, n$. Thus, the final simulated data is 
\ban 
Y_{i}  ( t_{i k} ) = \xi (t _{i k}) + \varepsilon (t_{i k}).
\ean

\item While this is supposed to be functional data, all we have really is a 
$n \times T$  table. Get the $T$-dimensional sample mean (minimizer of squared 
Euclidean norm), and median (ie, 
the minimizer of the Euclidean norm,  ie, Haldane's median, ie, the quantity that 
solves $\sum S(Y_{i} ; \theta) = 0$.) Also get your proposed weighted median.

\item Show that the weighted median and original median are robust, and the weighted 
median is closer than the median to the "population mean" 
\ban 
\sum_{j = 1}^{5} \beta_{j} PC_{j} (t), t = 1, \ldots, T.
\ean

\item Remember that although we are calling this FDA, this is really 12-dim data 
analysis. 
\end{enumerate}

\subsection{Subho's write-up}

Consider the general situation of estimating or testing for the location parameter of 
an elliptical distribution using weighted sign vectors. 
\bredbf
What is an elliptical distribution? In any scientific writing, for any technical term 
you must (a) either define it if it is brand new, (b) give the formula with proper 
citations if somebody has already defined it (in this case, there should be a definition 
in Fang's book, maybe in several other places), (c) prove every claim that you are making, 
or provide a reference from a trustworthy source.
\eredbf
For now the only condition 
we impose on these weights, say $w( \cdot)$, is that they need to be scalar-valued 
affine equivariant and square-integrable functions of the data, or in other words 
functions of the norm of the standardized random variable $\bfZ$.
\bredbf 
I am not sure if in FDA context, "affine equivariance" is defined. Find out. If not, then 
what you have above is a good definition. 
\eredbf
 In that sense $w(\bfX)$ can be equivalently written as $f(r)$, with $r = \| \bfZ \|$. 
 \bredbf 
 Are we always working with square integrable functions of the data? (to begin with, what 
 do you mean by that? Do you mean that the data is iid random variables all of which 
 have finite second moments? What is a standardized random variable? 
 \eredbf

 The simplest use of weighted signs here would be to construct a robust alternative to the 
 Hotelling's $T^2$ test \bredbf reference \eredbf using their sample mean vector and 
 covariance matrix. \bredbf what sort of robust alternative? provide details? always 
 provide 100\% exact mathematicalstatements \eredbf 
 Formally, 
 this means testing for $H_0: \bfmu = {\bf 0}_p$ vs. $H_1:\bfmu \neq {\bf 0}_p$  based 
 on the test statistic: \bredbf are you testing for the mean? Or a different parameter? 
 If it is the latter, then it's not the same test. Requires discussion and clarification \eredbf
%
$$ T_{n,w} = n \bar\bfX_w^T ( Cov (X_w))^{-1} \bar\bfX_w $$
%
with $\bar\bfX_w = \sum_{i=1}^n \bfX_{w,i}/n$ 
and $\bfX_{w,i} = w(\bfX_i ) \bfS (\bfX_i)$ for $i=1,2,...,n$. 
However, the following holds true for this weighted sign test:
%
\begin{Proposition}
Consider $n$ random variables $Z = (\bfZ_1,...,\bfZ_n)^T$ distributed independently and 
identically as $\mathcal{E}( \bfmu, kI_p, G); k \in \mathbb R$, and the class of hypothesis tests defined above. Then, given any $\alpha \in (0,1)$, local power 
at $\bfmu \neq {\bf 0}_p$ for the level-$\alpha$ test  based on $T_{n,w}$ is maximum when $w(\bfZ_1) = c$, a constant independent of $\bfZ_1$.
\end{Proposition}
%
\noindent This essentially means that power-wise the (unweighted) spatial sign test \citep{OjaBook10} is optimal in the given class of hypothesis tests when the data comes from a spherically symmetric distribution. Our simulations show that this empirically holds for non-spherical but elliptic distributions as well.

\subsection{The weighted spatial median}

Our weight functions are affine equivariant functions of the data, i.e. they are not affected by the population location parameter $\bfmu$. This ensures that there exists a unique solution of the following of optimization problem:
%
$$
\bfmu_w = \text{arg}\min_{\bfmu_0 \in \mathbb{R}^p} E ( w(\bfX) | \bfX - \bfmu_0 |)
$$
%
 This can be seen as a generalization of the Fermat-Weber location problem (which has the spatial median \citep{brown83, Chaudhuri96} as the solution) using data-dependent weights. We call its solution, $\bfmu_w$, the \textit{weighted spatial median} of $F$. In a sample setup it is estimated by iteratively solving the equation $\sum_{i=1}^n w(\bfX_i) \bfS (\bfX_i - \hat\bfmu_w)/n = {\bf 0}_p$.

The following theorem shows that the sample weighted spatial median $\hat\bfmu_w$ is a $\sqrt n$-consistent estimator of $\bfmu_w$, and gives its asymptotic distribution:

\begin{Theorem}
Let $A_w, B_w$ be two matrices, dependent on the weight function $w$ such that
%
$$
A_w = E \left[ \frac{w( \bfepsilon ) }{\| \bfepsilon \|} \left( 1 - \frac{\bfepsilon \bfepsilon^T}{\| \bfepsilon \|^2} \right) \right]; \quad B_w = E \left[ \frac{(w( \bfepsilon ))^2 \bfepsilon \bfepsilon^T}{\| \bfepsilon \|^2} \right]
$$
%
where $\bfepsilon \sim \mathcal E({\bf 0}_p, \Sigma, G)$. Then
%
\begin{equation}
\sqrt n (\hat\bfmu_w - \bfmu_w) \leadsto N_p ({\bf 0}_p, A_w^{-1} B_w A_w^{-1})
\end{equation}
\end{Theorem}
%

We provide a sketch of its proof in the supplementary material, which generalizes equivalent results for the spatial median \citep{OjaBook10}. Setting $w(\bfepsilon)=1$ above yields the asymptotic covariance matrix for the spatial median. Following this, the asymptotic relative efficiency (ARE) of $\bfmu_w$ corresponding to some non-uniform weight function with respect to the spatial median, say $\bfmu_s$ will be:
%
$$
ARE( \bfmu_w, \bfmu_s) = \left[ \frac{\text{det} (A^{-1} B A^{-1})}{\text{det} (A_w^{-1} B_w A_w^{-1})} \right]^{1/p}
$$
%
with $A = E [ 1/ \| \bfepsilon \| ( I_p - \bfepsilon \bfepsilon^T/ \| \bfepsilon \|^2 ) ]$ and $B = E [ \bfepsilon \bfepsilon^T/ \| \bfepsilon \|^2 ]$. This is further simplified under spherical symmetry:

\begin{Corollary}
For the spherical distribution $\mathcal{E}(\bfmu, kI_p, G); k \in \mathbb R, \bfmu \in \mathbb R^p$, we have
%
$$
ARE( \bfmu_w, \bfmu_s) = \frac{ \left[ E \left( \frac{f(r)}{r} \right) \right]^2}{Ef^2(r) \left[ E \left( \frac{1}{r} \right) \right]^2 }
$$
\end{Corollary}

\section{A high-dimensional test of location}

It is possible to take an alternative approach to the location testing problem by using the covariance-type U-statistic $C_{n,w} = \sum_{i=1}^n \sum_{j=1}^{i-1} \bfX_{w,i}^T \bfX_{w,j}$. This class of test statistics are especially attractive since they are readily generalized to cover high-dimensional situations, i.e. when $p > n$. The Chen and Qin (CQ) high-dimensional test of location for multivariate normal $\bfX_i$ \citep{ChenQin10} is a special case of this test that uses the statistic $C_n = \sum_{i=1}^n \sum_{j=1}^{i-1} \bfX_i^T \bfX_j$, and a recent paper (\citep{WangPengLi15}, from here on referred to as WPL test) shows that one can improve upon the power of the CQ test for non-gaussian elliptical distributions by using spatial signs $\bfS(\bfX_i)$ in place of the actual variables.

Given these, and some mild regularity conditions, the following holds for our generalized test statistic $C_{n,w}$ under $H_0$ as $n,p \rightarrow \infty$:
%
\begin{equation}\label{eqn:hdtest1}
\frac{C_{n,w}}{\sqrt{\frac{n(n-1)}{2} \text{Tr}(B_w^2)}} \leadsto N(0,1)
\end{equation}
%
and under contiguous alternatives $H_1: \bfmu = \bfmu_0$,
%
\begin{equation}\label{eqn:hdtest2}
\frac{C_{n,w} - \frac{n(n-1)}{2} \bfmu_0^T A_w^2 \bfmu_0 (1 + o(1)) }{\sqrt{\frac{n(n-1)}{2} \text{Tr}(B_w^2)}} \leadsto N(0,1)
\end{equation}
%
we provide the details behind deriving these two results in the supplementary material, which involve modified regularity conditions and sketches of proofs along the lines of \cite{WangPengLi15}.

Following this, the ARE of this test statistic with respect to its unweighted version, i.e. the WPL statistic, is expressed as:
%
$$
ARE(C_{n,w}, \text{WPL}; \bfmu_0) = \frac{\bfmu_0^T A_w^2 \bfmu_0}{\bfmu_0^T A^2 \bfmu_0} \sqrt\frac{\text{Tr}(B^2)}{\text{Tr}(B_w^2)} (1 + o(1))
$$
%
when $\Sigma = kI_p$, this again simplifies to $E^2(f(r)/r)/[E f^2(r). E^2(1/r)]$.

{\color{red}\bf (need to write the sketch in supplementary material) }

\section{SC: (older material) Testing for a location parameter}

Suppose $W \in \{ 0, 1 \}$, $Z \in \BR^{p}$ and $U \in \BR^{p}$ 
are independent  random variables, with 
$W$ having a Bernoulli distribution with parameter $\theta$, 
and each 
element of $U$ is an independent standard Cauchy random variable. 
The random vector $Z$ follows a mean-zero, variance matrix $\Sigma$ Normal distribution 
$Z \sim N_{p} (0, \Sigma)$. Our choices of $\Sigma$ in the details 
below include 
cases where $Cov (Z_{i}, Z_{j}) = \rho^{ | i - j|}$ for some $\rho \in [0, 1)$
which includes as special case the identity matrix $\BI_{p}$ corresponding to 
$\rho = 0$, and where $Cov (Z_{i}, Z_{j}) = \rho$.

We define $X = \mu + W Z + (1 - W) U$, and let the observed data be 
independent, identically distributed copies $X_{1}, \ldots, X_{n}$ copies of $X$. 
Note that the conditional distribution of $(X | W = 0)$ does not have finite moments.
Suppose our goal is to conduct inference on $\mu$, in particular, to test the hypothesis 
that $\mu = 0 \in \BR^{p}$. 
The sample mean and variance from the observed data $X_{1}, \ldots, X_{n}$, 
denoted respectively by $\bar{X} = n^{-1} \sum_{i = 1}^{n} X_{i}$ and
$S_{X} = (n - 1)^{-1} \sum_{i = 1}^{n} ( X_{i} - \bar{X} )( X_{i} - \bar{X} )^{T}$. 
These are 
well-defined quantities, even though their expectations may not exist.
Assume $p < n$, and thus $S_{x}$ is invertible.

One option for inference on $\mu$ is to ignore the heavy-tailed component of 
the data, and to use Hotteling's $T^{2}$ statistic given by 
$T^{2} = n (\bar{X} - \mu_{0})^{T} S_{X}^{-1} (\bar{X} - \mu_{0})$.

Other, more robust options, is to use the sign function exclusively, 
or any variety of data-depth functions exclusively. 
\bredbf suggest 2 or 3 depth functions used earlier for one-sample mean 
testing, provide references. List details on how these methods would be implemented,
just do what these references do
 \eredbf

Another alternative is to use the $\kappa$-transformation proposed in this paper. 
We begin with a robust estimator of $\mu$, for example the co-ordinatewise median, 
to use as $\mu_{X}$ and $\mu_{F}$. As $F$, we use the $p$-dimensional Normal 
distribution with mean $\mu_{F}$ and identity covariance matrix. One simple
peripherality function is $ || x - mu_{F} ||/ (1 + || x - mu_{F} ||)$, which 
is bounded above and below and achieves a minima at $\mu_{F}$. We take $\mu_{Y}$ 
to be $\mu$ also, and we may again use the co-ordinatewise median or a different 
statistic as its estimator like simple sample mean of $Y_{i} = \kappa (X_{i})$.
(Use the sample mean: we want to sell the idea that after the kappa transformation 
all standard classical methods like sample moments can be used).

We should be able to show that 
\begin{enumerate}

\item Classical methods like Hotteling is a disaster because of outliers.

\item Sign and depth-based methods lack power, which becomes quite bad when $\rho$ 
is considerably away from zero.

\end{enumerate}

\end{document}

\section{Testing for equality of two location parameters}

You know how to develop this. Do both the two sample and paired sample case.

Show simulation results like above.

\section{Develop a robust estimator of variance}
Simply do sample variance of the transformed variables $Y_{i} = \kappa (X_{i})$.

Show simulation results like above.

\section{Outlier detection}

Expand and generalize what you have in the paper already, where I think this is a small 
example. Refer to a standard method for multivariate outlier detection, and show that 
such a method used on $Y_{i} = \kappa (X_{i})$ works. I will send you some papers for 
referencing in this part.

\section{Robust principal component analysis}

This will be one of the bigger and 
 major sections of the paper, essentially copied and pasted
 from the previous version. Don't do anything here as of now.

\section{Robust Hawkins-regression}

Consider the $p + 1$ dimensional data $\{ 
(\mathbf{X}_{i} \ Y_{i})^{T} \in \BR^{p +1}, \ i =1, \ldots, n \}$. 
Do the robust PCA on this, and get all the eigenvectors. The typical eigenvector looks 
like $( \mathbf{\gamma} \ \gamma_{0} ) \in \BR^{p +1}$. Under the condition that the 
joint distribution of $(\mathbf{X} \ Y)^{T}$ is absolutely continuous, we have 
$\gamma_{0} \ne 0$ almost surely. We implement the following for any estimated 
eigenvector where $|\gamma_{0} | > \epsilon$ for some fixed $\epsilon > 0$: 
multiply the eigenvector by  $-\gamma_{0}^{-1}$, thus consider 
$( -\gamma_{0}^{-1} \mathbf{\gamma} \ -1 ) \in \BR^{p +1}$, which is a 
"Hawkins-regression". The  variance of this linear 
combination of variables is given by the corresponding eigenvalue times 
$\gamma_{0}^{-2}$, which may be small or large in the $p + 1$ eigenvectors, so the order 
in which eigenvalues/vectors appear in PCA may not be maintained here.

Also look up what Hawkins had originally proposed, and implement those.

Do a simple simulation, building on the simulation you have for the previous 
section.

\section{Robust inference with functional data}

This section is to show something beyond the $p \ll n$ setting. Both robust PCA 
and location testing are important problems for functional data. Abhirup can 
take charge of this part once we have everything else settled.

The main idea here is to use any decent location estimator to start with (some 
version of median is fine). Then we may test if the functional location for resting 
and active state are identical or not.

Also do functional PCA robust version, and then maybe project the data on the first few
principal components and then do the 2 sample (or paired sampel) testing again.

\section{An example with images}
\end{document}