### 2018-03-30 MS-new/7_Appendix.tex

\appendix
\section*{Appendix}
\numberwithin{equation}{section}
\section{\textbf{Form of $\BV_D(\BF)$}}
\label{section:appA}
First observe that for $\BF$ having covariance matrix $\Sigma = \Gamma\Lambda\Gamma^T$, we have
%
$$
\BV_D(\BF)  = (\Gamma \otimes \Gamma) \BV_D (\BF_\Lambda) (\Gamma \otimes \Gamma)^T,
$$
%
where $\BF_\Lambda$ has the same elliptic distribution as $\BF$, but with covariance matrix $\Lambda$. Now,
%
\begin{align*}
\BV_D (\BF_\Lambda) &= \BE \left[ \ve \left\{ \frac{(D^-_\bfZ (\bfz))^2 \Lambda^{1/2} \bfz\bfz^T \Lambda^{1/2}}{\bfz^T\Lambda\bfz} - \Lambda_D \right\} {\ve}^T \left\{ \frac{(D^-_\bfZ (\bfz))^2 \Lambda^{1/2} \bfz\bfz^T \Lambda^{1/2}}{\bfz^T\Lambda\bfz} - \Lambda_{D} \right\} \right]\\
&= \BE \left[ \ve \left\{ (D^-_\bfZ (\bfz))^2 \BS (\Lambda^{1/2}\bfz; \bf0) \right\} {\ve}^T \left\{ (D^-_\bfZ (\bfz))^2 \BS (\Lambda^{1/2}\bfz; \bf0) \right\} \right]
- \ve(\Lambda_{D}) {\ve}^T(\Lambda_{D})
\end{align*}

The matrix $\ve (\Lambda_{D}) {\ve}^T(\Lambda_{D})$ consists of elements $\lambda_{D,i} \lambda_{D,j}$ at $(i,j)^\text{th}$ position of the $(i,j)^\text{th}$ block, and 0 otherwise. These positions correspond to variance and covariance components of on-diagonal elements. For the expectation matrix, all its elements are of the form $ \BE [\sqrt{\lambda_a \lambda_b \lambda_c \lambda_d} z_a z_b z_c z_d . (D^-_\bfZ (\bfz))^4 / (\bfz^T \Lambda \bfz)^2]$, with $1 \leq a,b,c,d \leq p$. Since $(D^-_\bfZ (\bfz))^4 / (\bfz^T \Lambda \bfz)^2$ is even in $\bfz$, which has a sperically symmetric distribution, all such expectations will be 0 unless $a = b = c = d$, or they are pairwise equal. Following a similar derivation for spatial sign covariance matrices in \cite{magyar14}, we collect the non-zero elements and write the matrix of expectations:
%
$$
(\BI_{p^2} + \BK_{p,p}) \left\{ \sum_{a=1}^p \sum_{b=1}^p d_{ab} (\bfe_a \bfe_a^T \otimes  \bfe_b \bfe_b^T) - \sum_{a=1}^p d_{aa} (\bfe_a \bfe_a^T \otimes  \bfe_a \bfe_a^T) \right\} + \sum_{a=1}^p \sum_{b=1}^p d_{ab} (\bfe_a \bfe_b^T \otimes  \bfe_a \bfe_b^T) $$
%
where $\BI_k = (\bfe_1,...,\bfe_k), \BK_{m,n} = \sum_{i=1}^m \sum_{j=1}^n \BJ_{ij} \otimes \BJ_{ij}^T$ with $\BJ_{ij}$ the $m \times n$ matrix having 1 as $(i,j)^\text{th}$ element and 0 elsewhere, and $d_{mn} = \BE [ (D^-_\bfZ (\bfz))^4 (\BS_{mn}(\bfz; {\bf 0}))^2]; 1 \leq m,n \leq p$.

\paragraph{}Putting everything together, denote $ \tilde \BS(\BF_\Lambda) = \sum_{i=1}^n (\tilde D^{n-}_\bfZ (\bfz_i))^2 \BS (\Lambda^{1/2}\bfz_i; \hat \bfmu_n)/n $. Then the different types of elements in the matrix $\BV_D(F_\Lambda)$ are as given below ($1 \leq a,b,c,d \leq p$):

\begin{itemize}
\item Variance of on-diagonal elements
%
$$
A\BV( \sqrt n \tilde \BS_{aa}(\BF_\Lambda)) = \BE \left[ (D^-_\bfZ (\bfz))^4 (\BS_{aa} (\Lambda^{1/2} \bfz; {\bf 0}))^2 \right] - \lambda_{D,a}^2
$$

\item Variance of off-diagonal elements ($a \neq b$)
%
$$
A\BV( \sqrt n \tilde \BS_{ab}(\BF_\Lambda)) = \BE \left[ (D^-_\bfZ (\bfz))^4 (\BS_{ab} (\Lambda^{1/2} \bfz; {\bf 0}))^2 \right]
$$

\item Covariance of two on-diagonal elements ($a \neq b$)
%
$$
A\BC( \sqrt n \tilde \BS_{aa}(\BF_\Lambda), \sqrt n \tilde \BS_{bb}(\BF_\Lambda)) =
\BE \left[ (D^-_\bfZ (\bfz))^4 (\BS_{ab} (\Lambda^{1/2} \bfz; {\bf 0}))^2 \right] - \lambda_{D,a} \lambda_{D,b}
$$

\item Covariance of two off-diagonal elements ($a \neq b \neq c \neq d$)
%
$$
A\BC( \sqrt n \tilde \BS_{aa}(\BF_\Lambda), \sqrt n \tilde \BS_{bb}(\BF_\Lambda)) = 0
$$

\item Covariance of one off-diagonal and one on-diagonal element ($a \neq b \neq c$)
%
$$
A\BC( \sqrt n \tilde \BS_{ab}(\BF_\Lambda), \sqrt n \tilde \BS_{cc}(\BF_\Lambda)) = 0
$$
%
\end{itemize}

\section{Asymptotics of eigenvectors and eigenvalues}\label{section:appB}
The following result allows us to obtain asymptotic joint distributions of eigenvectors and eigenvalues of the sample DCM, provided we know the limiting distribution of the sample DCM itself:

\begin{Theorem}
\label{Theorem:decomp} \citep{taskinen12}
Let $\BF_\Lambda$ be an elliptical distribution with a diagonal covariance matrix $\Lambda$, and $\hat \BC$ be any positive definite symmetric $p \times p$ matrix such that at $\BF_\Lambda$ the limiting distribution of $\sqrt n \ve (\hat \BC - \Lambda)$ is a $p^2$-variate (singular) normal distribution with mean zero. Write the spectral decomposition of $\hat \BC$ as $\hat \BC = \hat \BP \hat \BL \hat \BP^T$. Then the limiting distributions of $\sqrt n vec(\hat \BP - \BI_p)$ and $\sqrt n vec(\hat\BL - \Lambda)$ are multivariate (singular) normal and
%
\begin{equation} \label{equation:decompEq}
\ve (\hat \BC - \Lambda) =
\left[ (\Lambda \otimes \BI_p) - (\BI_p \otimes \Lambda) \right]
\ve (\hat \BP - \BI_p) + \ve (\hat\Lambda - \BL) + o_P(n^{-1/2})
\end{equation}
\end{Theorem}

The first matrix picks only off-diagonal elements of the LHS and the second one only diagonal elements. We now use the above result and the form of $\BV_D(\BF)$ derived in Appendix~\ref{section:appA} to obtain limiting variance and covariances of eigenvalues and eigenvectors.

\begin{Corollary} \label{Corollary:eigendist}
Consider the sample DCM $ \tilde \BS = \sum_{i=1}^n (D^{n-}_\bfX (\bfx_i))^2 \BS (\bfx_i; \hat\bfmu_n)/n $ and its spectral decomposition $\tilde \BS = \hat\Gamma_D \hat\Lambda_D \hat\Gamma_D^T $. Then the matrices $\BG_D = \sqrt n (\hat\Gamma_D - \Gamma) $ and $\BL_D = \sqrt n (\hat\Lambda_D - \Lambda_{D}) $ have independent distributions. The random variable $\ve(\BG_D)$ asymptotically has a $p^2$-variate normal distribution with mean ${\bf 0}_{p^2}$, and the asymptotic variance and covariance of different columns of $\BG_D = (\bfg_1,...,\bfg_p)$ are as follows:
%
\begin{equation} \label{equation:DevEq}
A\BV(\bfg_i) = \sum_{k=1; k \neq i}^p \frac{\BE [ (D^-_\bfZ (\bfz))^4 (\BS_{ik}( \Lambda^{1/2} \bfz; {\bf 0}))^2 ]}
{(\lambda_{D,i} - \lambda_{D,k})^2} \bfgamma_k \bfgamma_k^T
\end{equation}
%
\begin{equation}
A\BC( \bfg_i, \bfg_j) = - \frac{\BE [ (D^-_\bfZ (\bfz))^4 (\BS_{ij}( \Lambda^{1/2} \bfz; {\bf 0}))^2 ]}
{(\lambda_{D,i} - \lambda_{D,j})^2} \bfgamma_i \bfgamma_j^T; \quad i \neq j
\end{equation}
%
The vector consisting of diagonal elements of $\BL_D$, say $\mathbf{\ell} = (\l_1,...,\l_p)^T$ asymptotically has a $p$-variate normal distribution with mean ${\bf 0}_p$ and variance-covariance elements:
%
\begin{eqnarray}
A\BV(l_i) &=&\BE [ (D^-_\bfZ (\bfz))^4 (\BS_{ii}( \Lambda^{1/2} \bfz; {\bf 0}))^2 ] - \lambda_{D,i}^2,\\
A\BC(l_i, l_j) &=& \BE [ (D^-_\bfZ (\bfz))^4 (\BS_{ij}( \Lambda^{1/2} \bfz; {\bf 0}))^2 ] - \lambda_{D,i} \lambda_{D,j}; \quad i \neq j.
\end{eqnarray}
%
\end{Corollary}

%\begin{proof}[Proof of Theorem \ref{Theorem:decomp}]
%See \cite{taskinen12}
%\end{proof}

\section{Proofs}
\label{section:appC}

\begin{proof}[Proof of Proposition \ref{proposition:SignTest}]
Under contiguous alternatives $H_0: \bfmu = \bfmu_0$, the weighted sign test statistic $T_{n,w}$ has mean $\BE (w(\bfZ) \bfS (\bfZ))$. For spherically symmetric $\bfZ$, $w(\bfZ)$ depends on $\bfZ$ only through its norm. Since $\| \bfZ \|$ and $\bfS(\bfZ)$ are independent, we get $\BE (w(\bfZ) \bfS (\bfZ)) = \BE w(\bfZ). \BE \bfS (\bfZ)$. A similar decomposition holds for $\BV (w(\bfZ) \bfS(\bfZ))$.

We can now simplify the approximate local power $\beta_{n,w}$ of a level-$\alpha$, $0 < \alpha < 1$, test based on $T_{n,w}$:
%
\begin{align*}
\beta_{n,w} &= K_p \left( \chi^2_{p,\alpha} + n (\BE (w(\bfZ) \bfS (\bfZ))^T
[ \BE (w^2(\bfZ) \bfS (\bfZ) \bfS (\bfZ)^T) ]^{-1} (\BE (w(\bfZ) \bfS (\bfZ)) \right)\\
&= K_p \left( \chi^2_{p,\alpha} + \frac{\BE^2 w(\bfZ)}{\BE w^2(\bfZ)}. \BE\bfS(\bfZ)^T [\BV \bfS(\bfZ)]^{-1} \BE \bfS(\bfZ) \right),
\end{align*}
%
where $K_p$ and $\chi^2_{p,\alpha}$ are distribution function and upper-$\alpha$ cutoff of a $\chi^2_p$ distribution, respectively. Since $\BE^2 w(\bfZ) \leq \BE w(\bfZ)$, the largest possible value of $\beta_{n,w}$ is $K_p ( \chi^2_{p,\alpha} + \BE \bfS(\bfZ)^T [\BV \bfS(\bfZ)]^{-1} \BE \bfS(\bfZ) )$, the approximate power of the unweighted sign test statistic. Equality is achieved when $w(\bfZ)$ is a constant independent of $\bfZ$.
\end{proof}

\begin{proof}[Sketch of proofs for equations \ref{eqn:hdtest1} and \ref{eqn:hdtest2}]

A first step to obtain asymptotic normality for the high-dimensional location test statistic $C_{n,w}$ is obtaining an equivalent result of Lemma 2.1 in \cite{WangPengLi15}:

\begin{Lemma}\label{Lemma:HDlemma21} Under the conditions

\noindent\textbf{(C1)}$\Tr(\Sigma^4) = o(\Tr^2(\Sigma^2)) $,

\noindent\textbf{(C2)}$\Tr^4(\Sigma) / \Tr^2(\Sigma^2) \exp[ - \Tr^2(\Sigma) / 128p \lambda^2_{\max}(\Sigma) ] = o(1)$
\vspace{1em}

\noindent when $H_0$ is true we have
%
\begin{eqnarray}
\BE[ (\bfepsilon_{w1}^T \bfepsilon_{w2})^4 ] &=& O(1) \BE^2[ (\bfepsilon_{w1}^T \bfepsilon_{w2})^2 ],\\
\BE[ (\bfepsilon_{w1}^T B_w \bfepsilon_{w1})^2 ] &=& O(1) \BE^2[ (\bfepsilon_{w1}^T B_w \bfepsilon_{w1})^2 ],\\
\BE[ (\bfepsilon_{w1}^T B_w \bfepsilon_{w2})^2 ] &=& o(1) \BE^2[ (\bfepsilon_{w1}^T B_w \bfepsilon_{w1})^2 ],
\end{eqnarray}
%
with $\bfepsilon \sim \mathcal E({\bf 0}_p, \Lambda, g)$ and $\bfepsilon_w = w(\bfepsilon) \bfS(\bfepsilon)$.
\end{Lemma}
%
A proof of this lemma is derived using results in section 3 of \cite{ElKaroui09}, noticing that any-scalar valued 1-Lipschitz function of $\bfepsilon_w$ is a $M_w$-Lipschitz function of $\bfS(\bfepsilon)$, with $M_w = \sup_\bfepsilon w(\bfepsilon)$. Same steps as in the proof of Theorem 2.2 in \cite{WangPengLi15} follow now, using the lemma above in place of Lemma 2.1 therein, to establish asymptotic normality of $C_{n,w}$ under $H_0$.

To derive the asymptotic distribution under contiguous alternatives we need the conditions (C3)-(C6) in \cite{WangPengLi15}, as well as slightly modified versions of Lemmas A.4 and A.5:

\begin{Lemma}
Given that condition (C3) holds, we have $\lambda_{\max} (\BB_w) \leq 2 \frac{\lambda_{\max}}{\text{Tr} (\Sigma)} (1+o(1))$.
\end{Lemma}

\begin{Lemma}
Define $\BD_w = \BE \left[ \frac{w^2(\bfepsilon)}{\| \bfepsilon \|^2} (\BI_p - \bfS(\bfepsilon) \bfS(\bfepsilon)^T )\right] $. Then $\lambda_{\max} (\BA_w) \leq \BE( w(\bfepsilon)/\| \bfepsilon \|)$ and $\lambda_{\max} (\BD_w) \leq \BE( w(\bfepsilon)/\| \bfepsilon \|)^2$. Further, if (C3) and (C4) hold then $\lambda_{\min} (\BA_w) \geq \BE( w(\bfepsilon)/\| \bfepsilon \|)(1+o(1))/\sqrt 3$.
\end{Lemma}
%
The proof now exactly follows steps in the proof of theorem 2.3 in \cite{WangPengLi15}, replacing vector signs by weighted signs, using the fact that $w(\bfepsilon)$ is bounded above by $M_w$ while applying conditions (C5)-(C6) and lemmas A.1, A.2, A.3, and finally using the above two lemmas in place of lemmas A.4 and A.5 respectively.
\end{proof}

\begin{proof}[Proof of Theorem  \ref{Theorem:covform}]
The proof follows directly from writing out the expression of $Cov ( \tilde \bfX)$:
%
\begin{eqnarray*}
\BV \tilde\bfX &=& \BE (\tilde\bfX \tilde\bfX^T) - \BE \tilde\bfX \BE \tilde\bfX^T\\
&=& \Gamma . \BE \left[ (D^-_\bfZ(\bfz))^2 \frac{\|\bfz\|^2}{\| \Lambda^{1/2}\bfz\|} \Lambda^{1/2} \bfS(\bfz) \bfS(\bfz)^T \Lambda^{1/2} \right] \Gamma^T - {\bf 0}_p {\bf 0}_p^T\\
&=& \Gamma . \BE \left[ (D^-_\bfZ(\bfz))^2 \frac{\Lambda^{1/2} \bfz \bfz^T \Lambda^{1/2}}{\bfz^T \Lambda \bfz} \right] \Gamma^T
\end{eqnarray*}
%
\end{proof}

\begin{proof}[Proof of Lemma \ref{Lemma:lemma1}]
For two positive definite matrices $\BA, \BB$, we denote by $ \BA > \BB$ that $\BA-\BB$ is positive definite. Also, denote
%
$$
\BS_n = \sqrt n \left[ \frac{1}{n} \sum_{i=1}^n \left| ( D^{n-} _\bfX (\bfx_i))^2  - (D^-_\bfX (\bfx_i))^2 \right| \BS( \bfx_i; \hat\bfmu_n) \right].
$$
%
%Consider now the quantity $\bft^T S_n \bft$, where $\bft \in \mathbb{R}^p$. Now applying Cauchy-Schwarz inequality we shall have
%%
%\begin{eqnarray}\label{eqn:AppEqn1}
%\frac{(\bft^T S_n \bft)^2}{n} & = & \left[ \frac{1}{n} \sum_{i=1}^n \left| (\tilde D^{n} _\bfX (\bfx_i))^2  - (D^-_\bfX (\bfx_i))^2 \right|^2  \bft^T SS(\bfx_i; \hat\bfmu_n) \bft \right]^2 \notag\\
%& \leq & \left[ \frac{1}{n} \sum_{i=1}^n \left| (\tilde D^{n} _\bfX (\bfx_i))^2  - (D^-_\bfX (\bfx_i))^2 \right|^2 \right] \left[ \frac{1}{n} \left( \sqrt n.\frac{1}{n} \sum_{i=1}^n  SS(\bfx, \hat\bfmu_n) \right) \right]
%\end{eqnarray}
%%
%with $\hat S_n = \sum_{i=1}^n SS(\bfx_i, \hat\bfmu_n)/ \sqrt n$.
Due to assumption (D1), given $\epsilon>0$ we can find $N \in \mathbb{N}$ such that
%
\begin{equation}
\label{equation:lemma1eq}
| (D^{n_1 -}_\bfX(\bfx_i))^2 - (D^-_\bfX(\bfx_i))^2 | < \epsilon
\end{equation}
%
for all $n_1 \geq N; i = 1,2,...,n_1$. This implies
%
\begin{eqnarray}
\BS_{n_1} &< & \epsilon \sqrt{n_1} \left[ \frac{1}{n_1} \sum_{i=1}^{n_1} \BS( \bfx_i; \hat\bfmu_{n_1}) \right]\notag\\
&=& \epsilon \sqrt{n_1} \left[ \frac{1}{n_1} \sum_{i=1}^{n_1} \left\{ \BS( \bfx_i; \hat\bfmu_{n_1}) - \BS( \bfx_i; \bfmu) \right\} + \frac{1}{n_1} \sum_{i=1}^{n_1} \BS( \bfx_i; \bfmu) \right]\label{eqn:lemma1eq2}
\end{eqnarray}

We now construct a sequence of positive definite matrices $\{a_k (B_k+C_k) : k \in \mathbb N\} $ so that
%
$$
a_k = \frac{1}{k}, \quad \BB_k = \sqrt{N_k} \left[ \frac{1}{N_k} \sum_{i=1}^{N_k} \left\{ \BS( \bfx_i; \hat\bfmu_{N_k}) - \BS(\bfx_i; \bfmu) \right\} \right],
\quad \BC_k = \sqrt{N_k} \left[ \frac{1}{N_k} \sum_{i=1}^{N_k} \BS(\bfx_i; \bfmu) \right]
$$
%
where $N_k \in \mathbb N$ gives the relation (\ref{equation:lemma1eq}) in place of $N$ when we take $\epsilon = 1/k$. Under conditions $ \BE \|\bfx - \bfmu\|^{-3/2} < \infty $ and $\sqrt n (\hat\bfmu_n - \bfmu) = O_P(1)$, the sample SCM with unknown location parameter $\hat\bfmu_n$ has the same asymptotic distribution as the SCM with known location $\bfmu$ \citep{durre14}, hence $\BB_k = o_P(1)$, thus $a_k (\BB_k + \BC_k) \stackrel{P}{\rightarrow} 0$.

Now \eqref{eqn:lemma1eq2} implies that for any $\epsilon_1 > 0$, $\BS_{N_k} > \epsilon_1 \Rightarrow a_k (\BB_k + \BC_k) > \epsilon_1$, which means $ P(\BS_{N_k} > \epsilon_1) < P(a_k (\BB_k + \BC_k) > \epsilon_1)$. Hence the subsequence $\{\BS_{N_k}\} \stackrel{P}{\rightarrow} 0$. Since the main sequence $\{\BS_k\}$ is bounded below by 0, this implies $\{\BS_k\} \stackrel{P}{\rightarrow} 0$. Finally, we have that
%
\begin{eqnarray}
\sqrt n \left[
\frac{1}{n} \sum_{i=1}^n ( D^{n-}_\bfX (\bfx_i))^2 \BS(\bfx_i; \hat\bfmu_n) -
\frac{1}{n} \sum_{i=1}^n (D^-_\bfX (\bfx_i))^2 \BS(\bfx_i; \bfmu) \right] &\leq & \hspace{1em} \notag\\
\BS_n +  \sqrt{n} \left[ \frac{1}{n} \sum_{i=1}^{n} \left\{ \BS(\bfx_i; \hat\bfmu_{n}) - \BS(\bfx_i; \bfmu) \right\} \right] &&
\end{eqnarray}
%
Since the second summand on the right hand side is $o_P(1)$ due to \cite{durre14} as mentioned before, we have the needed.
\end{proof}

\begin{proof}[Proof of Theorem \ref{Theorem:rootn}]
The quantity in the statement of the theorem can be broken down as:
%
\begin{eqnarray*}
\sqrt n \left[ \ve \left\{ \frac{1}{n} \sum_{i=1}^n ( D^{n-}_\bfX (\bfx_i))^2 \BS(\bfx_i; \hat\bfmu_n) \right\} - \ve \left\{ \frac{1}{n} \sum_{i=1}^n (D^-_\bfX (\bfx_i))^2 \BS( \bfx_i; \bfmu) \right\} \right] +\\
\sqrt n \left[ \ve \left\{ \frac{1}{n} \sum_{i=1}^n (D^-_\bfX (\bfx_i))^2 \BS( \bfx_i; \bfmu) \right\} - \BE \left[ \ve \left\{ (D^-_\bfX (\bfx))^2 \BS( \bfx; \bfmu) \right\} \right] \right]
\end{eqnarray*}
%
The first part goes to 0 in probability by Lemma \ref{Lemma:lemma1}, and applying Slutsky's theorem we get the required convergence.
\end{proof}

\begin{proof}[Proof of Theorem \ref{Thm:pluginSigma}]
We are going to prove the following:
%
\begin{enumerate}
\item $\| \hat\Gamma_D - \Gamma \|_F \stackrel{P}{\rightarrow} 0$, and

\item $\| \hat\Lambda - \Lambda \|_F \stackrel{P}{\rightarrow} 0$,
\end{enumerate}
%
as $n \rightarrow \infty$. For (1), we notice $\sqrt n \ve ( \hat\Gamma_D - \Gamma)$ asymptotically has a (singular) multivariate normal distribution following Corollary \ref{Corollary:eigendist}, so that $\| \hat\Gamma_D - \Gamma \|_F = O_P(1/\sqrt n)$ using Prokhorov's theorem.

It is now enough to prove convergence in probability of the individual eigenvalue estimates $\hat\lambda_i; i = 1,...,p$. For this, define estimates $\tilde\lambda_i$ as median-of-small-variances estimator of the \textit{true} score vectors $\Gamma^T \bfX$. For this we have
%
\begin{equation}\label{eqn:PluginSigmaProof1}
| \tilde\lambda_i - \lambda_i | \stackrel{P}{\rightarrow} 0
\end{equation}
%
using Theorem 3.1 of \cite{Minsker15}, with $\mu = \lambda_i$. Now $ \hat\lambda_i = \text{median}_j ( \BS ( \BX_{G_j} \hat\bfgamma_{D,i} ))$ and $\tilde\lambda_i = \text{median}_j ( \BS( \BX_{G_j} \bfgamma_i )) $, so that
%
\begin{eqnarray*}
| \hat\lambda_i - \tilde\lambda_i | &\leq & \text{median}_j \left[ \BS( \BX_{G_j} ( \hat\bfgamma_{D,i} - \bfgamma_i) ) \right]
\\ &\leq & \| \hat\bfgamma_{D,i} - \bfgamma_i \|^2 \text{median}_j  \left[ \text{Tr} ( \BS( \BX_{G_j}) ) \right]
\end{eqnarray*}
%
using Cauchy-Schwarz inequality. Combining the facts $ \| \hat\bfgamma_{D,i} - \bfgamma_i \| = O_P(1/\sqrt n)$ and $ \text{median}_j  [ \text{Tr} (\BS ( X_{G_j})) ] \stackrel{P}{\rightarrow} \text{Tr}(\Sigma)$ \citep{Minsker15} with (\ref{eqn:PluginSigmaProof1}), we get the needed.

\end{proof}

\begin{proof}[Proof of Corollary \ref{Corollary:eigendist}]
In spirit, this corollary is similar to Theorem 13.5.1 in \cite{anderson}. Due to the decomposition (\ref{equation:decompEq}) we have, for the distribution $F_\Lambda$, the following relation between any off-diagonal element of $\tilde \BS (\BF_\Lambda)$ and the corresponding element in the estimate of eigenvectors $\hat\Gamma_D (\BF_\Lambda)$:

$$ \sqrt n \hat\gamma_{D,ij} (\BF_\Lambda) = \sqrt n \frac{\tilde \BS_{ij} (\BF_\Lambda)}{\lambda_{D,i} - \lambda_{D,j}}; \quad i \neq j$$

So that for eigenvector estimates of the original $F$ we have
%
\begin{align}
\sqrt n (\hat\bfgamma_{D,i} - \bfgamma_i) &= \sqrt n \Gamma (\hat \bfgamma_{D,i}(\BF_\Lambda) - \bfe_i )\notag\\
&= \sqrt n \left[ \sum_{k=1; k \neq i}^p \hat \gamma_{D,ik}(\BF_\Lambda )\bfgamma_k + (\hat \gamma_{D,ii}(\BF_\Lambda) - 1)\bfgamma_i \right].\label{equation:app1}
\end{align}

By assumption we now have $\sqrt n (\hat \gamma_{D,ii}(\BF_\Lambda) - 1) =  o_P(1)$ and $A\BV( \sqrt n \tilde \BS_{ik}(\BF_\Lambda), \sqrt n \tilde \BS_{il}(\BF_\Lambda)) = 0$ for $k \neq l$, so the above equation implies
%
$$
A\BV(\bfg_i) = A\BV (\sqrt n (\hat\bfgamma_{D,i} - \bfgamma_i)) =
\sum_{k=1; k \neq i}^p \frac{A\BV(\sqrt n \tilde \BS_{ik}(\BF_\Lambda))}{(\lambda_{D,i} - \lambda_{D,k})^2} \bfgamma_k \bfgamma_k^T
$$

For the covariance terms, from (\ref{equation:app1}) we get, for $i \neq j$,
%
\begin{eqnarray*}
A\BV( \bfg_i, \bfg_j) &=& A\BV (\sqrt n (\hat\bfgamma_{D,i} - \bfgamma_i), \sqrt n (\hat\bfgamma_{D,j} - \bfgamma_j))\\
&=& A\BV \left( \sum_{k=1; k \neq i}^p \sqrt n \hat \gamma_{D,ik}(\BF_\Lambda)\bfgamma_k, \sum_{k=1; k \neq j}^p \sqrt n \hat \gamma_{D,jk}(\BF_\Lambda)\bfgamma_k \right)\\
&=& A\BV \left( \sqrt n \hat \gamma_{D,ij}(\BF_\Lambda)\bfgamma_j, \sqrt n \hat \gamma_{D,ji}(\BF_\Lambda)\bfgamma_i \right)\\
&=& - \frac{A\BV( \sqrt n \tilde \BS_{ij}(\BF_\Lambda))}{(\lambda_{D,i} - \lambda_{D,j})^2} \bfgamma_j \bfgamma_i^T.
\end{eqnarray*}

The exact forms given in the statement of the corollary now follows from the  Form of $\BV_D (\BF)$ in Appendix \ref{section:appA}.

\paragraph{}For the on-diagonal elements of $\tilde \BS(\BF_\Lambda)$, Theorem \ref{Theorem:decomp} gives us $ \sqrt n \hat\lambda_{D,i} ( \BF_\Lambda) = \sqrt n \tilde \BS_{ii}(\BF_\Lambda)$ for $i = 1,...,p$. Hence
%
\begin{eqnarray*}
A\BV(l_i) &=& A\BV(\sqrt n \hat\lambda_{D,i} - \sqrt n \lambda_{D,i})\\
&=& A\BV(\sqrt n \hat\lambda_{D,i} (\BF_\Lambda) - \sqrt n \lambda_{D,i}(\BF_\Lambda))\\
&=& A\BV(\sqrt n \BS_{ii}(\BF_\Lambda)).
\end{eqnarray*}

A similar derivation gives the expression for $A\BV(l_i,l_j); i \neq j$. Finally, since the asymptotic covariance between an on-diagonal and an off-diagonal element of $\tilde \BS( \BF_\Lambda)$, it follows that the elements of $\BG_D$ and diagonal elements of $\BL_D$ are independent.
\end{proof}

\section{Additional simulations}
\label{sec:appD}
Here we give tables of finite sample efficiencies of DCM and ADCMs for $p=2,3$.

\begin{table}[t]
\begin{scriptsize}
    \begin{tabular}{c|cc|ccc|ccc}
    \hline
    $F$ = Bivariate $t_5$    & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20                   & 0.80 & 0.83  & 0.95   & 0.95   & 0.89  & 1.00    & 0.96    & 0.89   \\
    $n$=50                   & 0.86 & 0.90  & 1.25   & 1.10   & 1.21  & 1.32    & 1.13    & 1.25   \\
    $n$=100                  & 1.02 & 1.04  & 1.58   & 1.20   & 1.54  & 1.67    & 1.24    & 1.63   \\
    $n$=300                  & 1.24 & 1.28  & 1.81   & 1.36   & 1.82  & 1.93    & 1.44    & 1.95   \\
    $n$=500                  & 1.25 & 1.29  & 1.80   & 1.33   & 1.84  & 1.91    & 1.39    & 1.97   \\ \hline
    $F$ = Bivariate $t_6$    & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20                   & 0.77 & 0.79  & 0.92   & 0.92   & 0.86  & 0.96    & 0.92    & 0.85   \\
    $n$=50                   & 0.76 & 0.78  & 1.11   & 1.00   & 1.08  & 1.17    & 1.03    & 1.13   \\
    $n$=100                  & 0.78 & 0.79  & 1.27   & 1.06   & 1.33  & 1.35    & 1.11    & 1.41   \\
    $n$=300                  & 0.88 & 0.91  & 1.29   & 1.09   & 1.35  & 1.38    & 1.15    & 1.45   \\
    $n$=500                  & 0.93 & 0.96  & 1.37   & 1.13   & 1.40  & 1.44    & 1.19    & 1.48   \\ \hline
    $F$ = Bivariate $t_{10}$ & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20                   & 0.70 & 0.72  & 0.83   & 0.84   & 0.77  & 0.89    & 0.87    & 0.79   \\
    $n$=50                   & 0.58 & 0.60  & 0.90   & 0.84   & 0.86  & 0.95    & 0.88    & 0.91   \\
    $n$=100                  & 0.57 & 0.59  & 0.92   & 0.87   & 0.97  & 0.98    & 0.90    & 1.03   \\
    $n$=300                  & 0.62 & 0.64  & 0.93   & 0.85   & 0.99  & 0.99    & 0.91    & 1.06   \\
    $n$=500                  & 0.62 & 0.65  & 0.93   & 0.86   & 1.00  & 1.00    & 0.92    & 1.08   \\ \hline
    $F$ = Bivariate $t_{15}$ & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20                   & 0.63 & 0.66  & 0.76   & 0.78   & 0.72  & 0.81    & 0.81    & 0.73   \\
    $n$=50                   & 0.52 & 0.52  & 0.79   & 0.75   & 0.80  & 0.84    & 0.79    & 0.85   \\
    $n$=100                  & 0.51 & 0.52  & 0.83   & 0.77   & 0.88  & 0.88    & 0.81    & 0.94   \\
    $n$=300                  & 0.55 & 0.56  & 0.84   & 0.79   & 0.91  & 0.89    & 0.84    & 0.98   \\
    $n$=500                  & 0.56 & 0.59  & 0.85   & 0.80   & 0.93  & 0.91    & 0.86    & 0.99   \\ \hline
    $F$ = Bivariate $t_{25}$ & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20                   & 0.63 & 0.65  & 0.77   & 0.79   & 0.74  & 0.80    & 0.81    & 0.74   \\
    $n$=50                   & 0.49 & 0.50  & 0.73   & 0.71   & 0.76  & 0.78    & 0.75    & 0.80   \\
    $n$=100                  & 0.45 & 0.46  & 0.73   & 0.69   & 0.81  & 0.78    & 0.73    & 0.87   \\
    $n$=300                  & 0.51 & 0.52  & 0.78   & 0.75   & 0.87  & 0.83    & 0.79    & 0.94   \\
    $n$=500                  & 0.53 & 0.55  & 0.79   & 0.75   & 0.87  & 0.84    & 0.80    & 0.94   \\ \hline
    $F$ = BVN                & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20                   & 0.56 & 0.60  & 0.69   & 0.71   & 0.67  & 0.73    & 0.74    & 0.68   \\
    $n$=50                   & 0.42 & 0.43  & 0.66   & 0.66   & 0.70  & 0.71    & 0.69    & 0.75   \\
    $n$=100                  & 0.42 & 0.43  & 0.69   & 0.66   & 0.77  & 0.74    & 0.71    & 0.83   \\
    $n$=300                  & 0.47 & 0.49  & 0.71   & 0.69   & 0.82  & 0.76    & 0.73    & 0.88   \\
    $n$=500                  & 0.48 & 0.50  & 0.73   & 0.71   & 0.83  & 0.78    & 0.76    & 0.89   \\ \hline
    \end{tabular}
\end{scriptsize}
\caption{Finite sample efficiencies of several scatter matrices: $p=2$}
\label{table:FSEtable2}
\end{table}

\begin{table}[t]
\begin{scriptsize}
   \begin{tabular}{c|cc|ccc|ccc}
    \hline
    3-variate $t_5$    & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20             & 0.96 & 0.97  & 1.06   & 1.03   & 0.99  & 1.07    & 1.06    & 0.97   \\
    $n$=50             & 1.07 & 1.08  & 1.28   & 1.20   & 1.18  & 1.33    & 1.23    & 1.20   \\
    $n$=100            & 1.12 & 1.15  & 1.49   & 1.31   & 1.40  & 1.57    & 1.38    & 1.48   \\
    $n$=300            & 1.49 & 1.54  & 2.09   & 1.82   & 2.07  & 2.19    & 1.93    & 2.18   \\
    $n$=500            & 1.60 & 1.66  & 2.18   & 1.87   & 2.21  & 2.27    & 1.95    & 2.30   \\ \hline
    3-variate $t_6$    & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20             & 0.90 & 0.92  & 1.00   & 0.99   & 0.95  & 1.02    & 1.01    & 0.94   \\
    $n$=50             & 0.95 & 0.96  & 1.16   & 1.09   & 1.09  & 1.21    & 1.14    & 1.11   \\
    $n$=100            & 0.98 & 0.99  & 1.32   & 1.22   & 1.25  & 1.38    & 1.27    & 1.29   \\
    $n$=300            & 1.10 & 1.14  & 1.57   & 1.40   & 1.58  & 1.62    & 1.47    & 1.64   \\
    $n$=500            & 1.17 & 1.20  & 1.57   & 1.43   & 1.60  & 1.63    & 1.51    & 1.67   \\ \hline
    3-variate $t_{10}$ & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20             & 0.87 & 0.88  & 0.95   & 0.94   & 0.90  & 0.97    & 0.98    & 0.89   \\
    $n$=50             & 0.77 & 0.79  & 0.96   & 0.92   & 0.94  & 0.99    & 0.96    & 0.95   \\
    $n$=100            & 0.75 & 0.76  & 1.02   & 0.95   & 1.01  & 1.06    & 1.00    & 1.05   \\
    $n$=300            & 0.73 & 0.75  & 1.03   & 0.98   & 1.10  & 1.08    & 1.03    & 1.15   \\
    $n$=500            & 0.73 & 0.76  & 1.02   & 0.98   & 1.09  & 1.06    & 1.02    & 1.14   \\ \hline
    3-variate $t_{15}$ & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20             & 0.84 & 0.86  & 0.92   & 0.92   & 0.89  & 0.94    & 0.94    & 0.87   \\
    $n$=50             & 0.75 & 0.76  & 0.92   & 0.90   & 0.90  & 0.96    & 0.94    & 0.93   \\
    $n$=100            & 0.66 & 0.67  & 0.91   & 0.87   & 0.95  & 0.96    & 0.92    & 1.00   \\
    $n$=300            & 0.61 & 0.64  & 0.90   & 0.87   & 1.00  & 0.93    & 0.91    & 1.04   \\
    $n$=500            & 0.65 & 0.67  & 0.89   & 0.87   & 0.99  & 0.93    & 0.91    & 1.03   \\ \hline
    3-variate $t_{25}$ & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20             & 0.78 & 0.79  & 0.87   & 0.89   & 0.87  & 0.89    & 0.92    & 0.86   \\
    $n$=50             & 0.70 & 0.71  & 0.88   & 0.86   & 0.88  & 0.91    & 0.90    & 0.90   \\
    $n$=100            & 0.61 & 0.63  & 0.86   & 0.83   & 0.89  & 0.90    & 0.88    & 0.94   \\
    $n$=300            & 0.58 & 0.59  & 0.83   & 0.80   & 0.92  & 0.87    & 0.85    & 0.98   \\
    $n$=500            & 0.62 & 0.64  & 0.83   & 0.82   & 0.94  & 0.88    & 0.87    & 0.99   \\ \hline
    3-variate Normal   & SCM  & Tyler & DCM-H & DCM-M & DCM-P & ADCM-H & ADCM-M & ADCM-P \\ \hline
    $n$=20             & 0.76 & 0.78  & 0.85   & 0.87   & 0.84  & 0.87    & 0.90    & 0.83   \\
    $n$=50             & 0.66 & 0.67  & 0.82   & 0.81   & 0.84  & 0.86    & 0.86    & 0.86   \\
    $n$=100            & 0.56 & 0.58  & 0.77   & 0.75   & 0.83  & 0.82    & 0.79    & 0.87   \\
    $n$=300            & 0.53 & 0.55  & 0.75   & 0.74   & 0.85  & 0.79    & 0.78    & 0.90   \\
    $n$=500            & 0.56 & 0.58  & 0.76   & 0.76   & 0.87  & 0.80    & 0.80    & 0.92   \\ \hline
    \end{tabular}
\end{scriptsize}
\caption{Finite sample efficiencies of several scatter matrices: $p=3$}
\label{table:FSEtable3}
\end{table}