### 2018-03-30 MS-new/Signed-peripherality-draft.tex

\documentclass[11pt,letterpaper]{article}
\pdfoutput=1

\usepackage{mycommands,amssymb,amsmath,amsthm,color,pagesize,outlines,cite,subfigure,epsfig}
\usepackage[small]{caption}
\usepackage{hyperref} % for linking references
\hypersetup{colorlinks = true, citecolor = blue, urlcolor = blue}

\usepackage{stackrel}

\usepackage[round]{natbib}

% for algorithm
%\usepackage[noend]{algpseudocode}
%\usepackage{algorithm}

% DON'T change margins - should be 1 inch all around.
\addtolength{\evensidemargin}{-.5in}
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\textwidth}{0.9in}
\addtolength{\textheight}{0.9in}
\addtolength{\topmargin}{-.4in}

%% measurements for 1 inch margin
%\addtolength{\oddsidemargin}{-.875in}
%\addtolength{\evensidemargin}{-.875in}
%\addtolength{\textwidth}{1.75in}
%\addtolength{\topmargin}{-.875in}
%\addtolength{\textheight}{1.75in}

%\pagestyle{myheadings}
%\markboth{}{\underline{{\bf Notes: (do not circulate)} \hspace{4.5cm} {\sc  Ansu Chatterjee} \hspace{0.25cm}}}

\DeclareMathOperator*{\ve}{vec}
\DeclareMathOperator*{\diag}{diag }
\DeclareMathOperator*{\supp}{supp }
\DeclareMathOperator*{\Tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\Th}{^{\text{th}}}

\makeatletter
\newcommand{\opnorm}{\@ifstar\@opnorms\@opnorm}
\newcommand{\@opnorms}[1]{%
  \left|\mkern-1.5mu\left|\mkern-1.5mu\left|
   #1
  \right|\mkern-1.5mu\right|\mkern-1.5mu\right|
}
\newcommand{\@opnorm}[2][]{%
  \mathopen{#1|\mkern-1.5mu#1|\mkern-1.5mu#1|}
  #2
  \mathclose{#1|\mkern-1.5mu#1|\mkern-1.5mu#1|}
}
\makeatother

%% Appendix theorem counter
\usepackage{chngcntr}
\usepackage{apptools}
\AtAppendix{\counterwithin{Theorem}{section}}
\numberwithin{equation}{section}

\begin{document}

\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Corollary}[Theorem]{Corollary}
\newtheorem{Proposition}[Theorem]{Proposition}
\newtheorem{Conjecture}[Theorem]{Conjecture}
\theoremstyle{definition} \newtheorem{Definition}[Theorem]{Definition}

\title{Signed Peripherality Functions for Multivariate Statistical Inference}
\date{}
\author{Subhabrata Majumdar, Snigdhansu Chatterjee}
\maketitle

Abstract:
We introduce a notion of multivariate ranking based on the composition of the multivariate sign function and data-dependent weights that are increasing or decreasing functions of a point-to-distribution distance measure in multivariate space, which we call the peripherality function. Using this weighted sign vector, we propose estimation and testing methods for the mean vector of an elliptical distribution that improve upon corresponding spatial sign-based procedures. When bounded and monotonically decreasing transformations on data depth functions are used as weights, the rank vectors can be used as proxies of the original data to estimate different components of the covariance matrix in the data-generating elliptical distribution: its eigenvectors, eigenvalues and the covariance matrix itself. Wimplement our proposed methods to perform robust Sufficient Dimension Reduction \citep{AdragniCook09} and functional outlier detection. Throughout the paper, several simulation studies and data examples outline the robustness and efficiency of our methods.
\vspace{.5cm}

Keywords:
Multivariate ranking, data depth, high-dimensional location testing, principal component analysis, sufficient dimension reduction, robust PCA, outlier detection

\newpage
\input{1_Introduction}
\input{2_Location}
\input{3_Scale}
\input{4_SDR}
\input{5_Functional}
\input{6_Conclusion}

\input{7_Appendix}

%\newpage
%\input{misc}

\bibliographystyle{apalike}
\bibliography{scatterbib}

\end{document}