### 2019-09-13 EJSSubmission/WeightedSign_EJS-supp.tex
supple done

\documentclass[ejs]{imsart}

\RequirePackage[OT1]{fontenc}
%\RequirePackage{amsthm,amsmath}
\RequirePackage[numbers]{natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{amssymb,amsmath,amsthm,color,outlines,subfigure,comment}
\usepackage[small]{caption}
%\usepackage{hyperref} % for linking references
%\hypersetup{colorlinks = true, citecolor = blue, urlcolor = blue}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{stackrel}
\usepackage{natbib}
\usepackage{soul}
%\usepackage[bordercolor=white,backgroundcolor=gray!30,linecolor=black,colorinlistoftodos]{todonotes}
%\newcommand{\rework}[1]{\todo[color=yellow,inline]{#1}}

% will be filled by editor:
\doi{10.1214/154957804100000000}
\pubyear{0000}
\volume{0}
\firstpage{0}
\lastpage{0}
%\arxiv{}

%%%settings
%\pubyear{2006}
%\volume{0}
%\issue{0}
%\firstpage{1}
%\lastpage{8}

%%%%%% put your definitions there:
\startlocaldefs
%\numberwithin{equation}{section}
%\theoremstyle{plain}
%\newtheorem{thm}{Theorem}[section]

\newcommand{\BA}{{\mathbb{A}}}
\newcommand{\BB}{{\mathbb{B}}}
\newcommand{\BC}{{\mathbb{C}}}
\newcommand{\BE}{{\mathbb{E}}}
\newcommand{\BD}{{\mathbb{D}}}
\newcommand{\BF}{{\mathbb{F}}}
\newcommand{\BG}{{\mathbb{G}}}
\newcommand{\BH}{{\mathbb{H}}}
\newcommand{\BI}{{\mathbb{I}}}
\newcommand{\BJ}{{\mathbb{J}}}
\newcommand{\BK}{{\mathbb{K}}}
\newcommand{\BL}{{\mathbb{L}}}
\newcommand{\BP}{{\mathbb{P}}}
\newcommand{\BR}{{\mathbb{R}}}
\newcommand{\BS}{{\mathbb{S}}}
\newcommand{\BT}{{\mathbb{T}}}
\newcommand{\BU}{{\mathbb{U}}}
\newcommand{\BV}{{\mathbb{V}}}
\newcommand{\BW}{{\mathbb{W}}}
\newcommand{\BX}{{\mathbb{X}}}
\newcommand{\BY}{{\mathbb{Y}}}
\newcommand{\BZ}{{\mathbb{Z}}}

\newcommand{\BCOV}{{\mathbb{COV}}}

\newcommand{\colrit}{\color{red} \it}
\newcommand{\colrbf}{\color{red} \bf}
\newcommand{\colbit}{\color{blue} \it}
\newcommand{\colbbf}{\color{blue} \bf}
\newcommand{\colmit}{\colm \it}
\newcommand{\colmbf}{\colm \bf}

\newcommand{\bfOne}{\bf 1}
\newcommand{\iid}{\stackrel{i.i.d.}{\sim}}
\newcommand{\raro}{\rightarrow}

\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cC}{\mathcal{C}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cE}{\mathcal{E}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cI}{\mathcal{I}}
\newcommand{\cM}{\mathcal{M}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cY}{\mathcal{Y}}

\def\baq#1\eaq{\begin{align}#1\end{align}}
\def\ban#1\ean{\begin{align*}#1\end{align*}}
\newcommand{\draro}{ \stackrel{{\mathcal D}}{\Rightarrow} }

\def\bredbf#1\eredbf{{\color{red}{\bf ???? #1 ????}}}
\def\BeginRedComment#1\EndRedComment{({\color{red}{\bf Comment:} {\it #1}}) }
\def\BeginBlueComment#1\EndBlueComment{({\color{blue}{\bf Comment:} {\it #1}}) }

\DeclareMathOperator*{\ve}{vec}
\DeclareMathOperator*{\diag}{diag }
\DeclareMathOperator*{\supp}{supp }
\DeclareMathOperator*{\Tr}{Tr}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\Th}{^{\text{th}}}

\makeatletter
\newcommand{\opnorm}{\@ifstar\@opnorms\@opnorm}
\newcommand{\@opnorms}[1]{%
  \left|\mkern-1.5mu\left|\mkern-1.5mu\left|
   #1
  \right|\mkern-1.5mu\right|\mkern-1.5mu\right|
}
\newcommand{\@opnorm}[2][]{%
  \mathopen{#1|\mkern-1.5mu#1|\mkern-1.5mu#1|}
  #2
  \mathclose{#1|\mkern-1.5mu#1|\mkern-1.5mu#1|}
}
\makeatother

\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Corollary}[Theorem]{Corollary}
\newtheorem{Proposition}[Theorem]{Proposition}
\newtheorem{Conjecture}[Theorem]{Conjecture}
\theoremstyle{definition} \newtheorem{Definition}[Theorem]{Definition}

\endlocaldefs

\begin{document}

\begin{frontmatter}
\title{On Weighted Multivariate Sign Functions: Supplementary Material}
\runtitle{Weighted Sign Functions}
%\thankstext{T1}{Footnote to the title with the `thankstext' command.}

\begin{aug}
\author{\fnms{Subhabrata} \snm{Majumdar}\thanksref{t1}\ead[label=e1]{subho@research.att.com}}
\and
\author{\fnms{Snigdhansu} \snm{Chatterjee}\thanksref{t3}\ead[label=e2]{chatt019@umn.edu}}

\address{School of Statistics,\\
University of Minnesota,\\
313 Ford Hall,\\
224 Church Street S.E., \\
Minneapolis 55455, USA\\
%usually few lines long\\
\printead{e1,e2}}

%\author{\fnms{Third} \snm{Author}
%\ead[label=e3]{third@somewhere.com}
%\ead[label=u1,url]{www.foo.com}}
%
%\address{Address of the Third author\\
%usually few lines long\\
%usually few lines long\\
%\printead{e3}\\
%\printead{u1}}

\thankstext{t1}{Currently at AT\&T Labs Research}
%\thankstext{t2}{First supporter of the project}
\thankstext{t3}{Corresponding author}
\runauthor{Majumdar and Chatterjee}

\affiliation{University of Minnesota and University of Minnesota}

\end{aug}

\end{frontmatter}

\appendix
\renewcommand{\thesection}{\Alph{section}}
\section{Form of $V_W$}\label{sec:appA}
First observe that for $F$ having covariance matrix $\Sigma = \Gamma\Lambda\Gamma^T$,
%
\begin{align}\label{eqn:VWeqn}
V_W  = (\Gamma \otimes \Gamma) V_{W, \Lambda} (\Gamma \otimes \Gamma)^T
\end{align}
%
where $V_{W, \Lambda}$ is the covariance matrix of $\BF_\Lambda$, the elliptic distribution with mean $\mu$ and covariance matrix $\Lambda$. Now,
%
\begin{eqnarray*}
V_{W, \Lambda} &=& \BE \left[ \ve \left\{ \frac{W^2 (Z, \BF_Z) \Lambda^{1/2} ZZ^T \Lambda^{1/2} }{ Z^T \Lambda Z} - \tilde \Lambda \right\} {\ve}^T \left\{ \frac{ W^2(Z, \BF_Z) \Lambda^{1/2} ZZ^T \Lambda^{1/2} }{ Z^T \Lambda Z} - \tilde \Lambda \right\} \right]\\
&=& \BE \left[ \ve \left\{ W^2 (Z, \BF_Z) \BS( \Lambda^{1/2} z; {\bf 0} ) \right\} {\ve}^T \left\{ W^2(Z, \BF_Z) \BS( \Lambda^{1/2}Z; \bf0) \right\} \right]\\
&& - \quad \ve( \tilde \Lambda) {\ve}^T( \tilde \Lambda )
\end{eqnarray*}

The matrix $\ve( \tilde\Lambda) \ve^T(\tilde\Lambda)$ consists of elements $\lambda_i\lambda_j$ at $(i,j)^\text{th}$ position of the $(i,j)^\text{th}$ block, and 0 otherwise. These positions correspond to variance and covariance components of on-diagonal elements. For the expectation matrix, all its elements are of the form $ \BE[ \sqrt{\lambda_a \lambda_b \lambda_c \lambda_d} Z_a Z_b Z_c Z_d . W^4 (Z, \BF_Z) / (Z^T \Lambda Z)^2]$, with $1 \leq a,b,c,d \leq p$. Since $W^4 (Z, \BF_Z) / (Z^T \Lambda Z)^2$ is even in $Z$, which has a spherically symmetric distribution, all such expectations will be 0 unless $a,b,c,d$ are all equal or pairwise equal. Following a similar derivation for spatial sign covariance matrices in \cite{ref:Biometrika14673_MagyarTyler}, we collect the non-zero elements and write the matrix of expectations:
%
$$ (\BI_{p^2} + \BK_{p,p}) \left\{ \sum_{a=1}^p \sum_{b=1}^p \tilde\gamma_{ab} (e_a e_a^T \otimes  e_b e_b^T) - \sum_{a=1}^p \tilde\gamma_{aa} (e_a e_a^T \otimes  e_a e_a^T) \right\} + \sum_{a=1}^p \sum_{b=1}^p \tilde\gamma_{ab} (e_a e_b^T \otimes  e_a e_b^T) $$
%
where $\BI_k = (e_1,...,e_k), \BK_{m,n} = \sum_{i=1}^m \sum_{j=1}^n \BJ_{ij} \otimes \BJ_{ij}^T$ with $\BJ_{ij} \in \BR^{m \times n}$ having 1 as $(i,j)$-th element and 0 elsewhere, and $\tilde\gamma_{mn} = \BE[ \lambda_m \lambda_n Z_m^2 Z_n ^2 . W^4 (Z, \BF_Z) / (Z^T \Lambda Z)^2]; 1 \leq m,n \leq p$.

\paragraph{}Putting everything together, denote by $\hat {\tilde \Lambda}$ the sample version of $\tilde \Lambda$, the weighted covariance matrix obtained from $\BF_\Lambda$, i.e. $\hat {\tilde \Lambda} = \sum_{i=1}^n W_n^2 (Z_i, \BF_{Z,n}) \BS( \Lambda^{1/2} Z_i; \hat \mu_n)/n $. Then the different types of elements in the matrix $\hat {\tilde \Lambda}$ are as given below ($1 \leq a,b,c,d \leq p$):

\begin{itemize}
\item Variance of on-diagonal elements
%
$$ A\BV( \sqrt n \hat {\tilde \Lambda} (a,a) ) = \BE \left[ \frac{ W^4(Z, \BF_Z) \lambda_a^2 Z_a^4}{(Z^T \Lambda Z)^2} \right] - \tilde\lambda_a^2 $$

\item Variance of off-diagonal elements ($a \neq b$)
%
$$ A\BV( \sqrt n \hat {\tilde \Lambda} (a,b) ) = \BE \left[ \frac{W^4 (Z, \BF_Z) \lambda_a \lambda_b Z_a^2 Z_b^2}{(Z^T \Lambda Z)^2} \right] $$

\item Covariance of two on-diagonal elements ($a \neq b$)
%
$$ A\BV(\sqrt n \hat {\tilde \Lambda} (a,a), \sqrt n \hat {\tilde \Lambda} (b,b) )
= \BE \left[ \frac{W^4 (Z, \BF_Z) \lambda_a \lambda_b Z_a^2 Z_b^2}{(Z^T \Lambda Z)^2} \right] - \tilde\lambda_{a} \tilde\lambda_{b} $$

\item Covariance of two off-diagonal elements ($a \neq b, c \neq d$)
%
$$ A\BV(\sqrt n \hat {\tilde \Lambda} (a,b), \sqrt n \hat {\tilde \Lambda} (c,d) ) = 0 $$

\item Covariance of one off-diagonal and one on-diagonal element ($a \neq b \neq c$)
%
$$ A\BV(\sqrt n \hat {\tilde \Lambda} (a,b), \sqrt n \hat {\tilde \Lambda} (c,c) ) = 0 $$
\end{itemize}

\noindent The above give all the elements of $V_{W,\Lambda}$. We plug these in \eqref{eqn:VWeqn} to recover $V_W$.

\section{Proofs}\label{section:appB}

\begin{proof}[Proof sketch of Corollary 2.4]
%A first step to obtain asymptotic normality for the high-dimensional location test statistic $C_{n,w}$ is obtaining an equivalent result of Lemma 2.1 in \cite{ref:JASA151658_WangPengLi}:
%
%\begin{Lemma}\label{Lemma:HDlemma21} Consider a random variable $\epsilon$ having an elliptical distribution with mean ${\bf 0}$ and covariance matrix $\Lambda$. Then, under the conditions 1-2 in Corollary 2.4, we have
%%
%\begin{eqnarray}
%\BE[ (\epsilon_{W1}^T \epsilon_{W2})^4 ] &=& O(1) \BE^2[ (\epsilon_{W1}^T \epsilon_{W2})^2 ]\\
%\BE[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W1})^2 ] &=& O(1) \BE^2[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W1})^2 ]\\
%\BE[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W2})^2 ] &=& o(1) \BE^2[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W1})^2 ]
%\end{eqnarray}
%%
%with $\epsilon_W = W(\epsilon, \BF_\epsilon) S(\epsilon)$ where $\BF_\epsilon$ is the distribution of $\epsilon$.
%\end{Lemma}
%%
%A proof of this lemma is derived using results in section 3 of \cite{ElKaroui09}, noticing that any-scalar valued 1-Lipschitz function of $\epsilon_W$ is a $W_\max$-Lipschitz function of $S(\epsilon)$.

%To derive the asymptotic distribution under contiguous alternatives we need the conditions (C3)-(C6) in \cite{ref:JASA151658_WangPengLi}, as well as
We start with slightly modified versions of Lemmas A.4 and A.5 in \cite{ref:JASA151658_WangPengLi}:

\begin{Lemma}\label{lemma:lemmaB1}
Given that condition 1 in Corollary 2.4 holds, we have $\lambda_{\max} (\Psi_{1W}) \leq 2 W_{\max} \frac{\lambda_{\max} (\Sigma) }{\text{Tr} (\Sigma)} (1+o(1))$.
\end{Lemma}

\begin{Lemma}\label{lemma:lemmaB2}
Define $\Psi_{3W} = \BE \left[ \frac{W^2(\epsilon, \BF_\epsilon)}{| \epsilon |^2} (\BI_p - S(\epsilon) S(\epsilon)^T )\right] $. Then $\lambda_{\max} (\Psi_{2W}) \leq W_{\max} \BE(| \epsilon |^{-1})$ and $\lambda_{\max} (\Psi_{3W}) \leq (W_{\max})^2 \BE( | \epsilon|^{-2})$. Further, if conditions 1 and 2 of Corollary 2.4 hold then $\lambda_{\min} (\Psi_{2W}) \geq \BE( W(\epsilon, \BF_\epsilon)/ | \epsilon |)(1+o(1))/\sqrt 3$.
\end{Lemma}
%
The lemmas can be proved using similar steps as the proofs of the original lemmas in \cite{ref:JASA151658_WangPengLi} and using the upper bound on the weight function. Corollary 2.4 is now proved by applying Corollary 2.2, plugging in upper bound of $\lambda_{\max} (\Psi_{1W})$ from Lemma~\ref{lemma:lemmaB1} and lower bound of $\lambda_{\max} (\Psi_{1W})$ from Lemma~\ref{lemma:lemmaB2} into the ARE expression.
\end{proof}

\begin{proof}[Proof of Theorem 3.4]
We suppose $G_n = (g_1, \ldots, g_p), L_n = \diag(l_1, \ldots, l_p)$. In spirit, this corollary is similar to Theorem 13.5.1 in \cite{ref:AndersonBook09}. We start with the following result, due to \citep{ref:SPL12765_Taskinenetal}, allows us to obtain asymptotic joint distributions of eigenvectors and eigenvalues of $\hat{\tilde \Sigma}$, provided we know the limiting distribution of $\hat{\tilde \Sigma}$ itself:

\begin{Theorem} \label{Theorem:decomp} 
Let $\BF_\Lambda$ be defined as before, and $\hat C$ be any positive definite symmetric $p \times p$ matrix such that at $F_\Lambda$ the limiting distribution of $\sqrt n \ve(\hat C - \Lambda)$ is a $p^2$-variate (singular) normal distribution with mean zero. Write the spectral decomposition of $\hat C$ as $\hat C = \hat P \hat\Lambda \hat P^T$. Then the limiting distributions of $\sqrt n \ve(\hat P - \BI_p)$ and $\sqrt n \ve( \hat\Lambda - \Lambda)$ are multivariate (singular) normal and
%
\begin{equation} \label{equation:decompEq}
\sqrt n \ve (\hat C - \Lambda)  = \left[ (\Lambda \otimes \BI_p) - (\BI_p \otimes \Lambda) \right] \sqrt n \ve (\hat P - \BI_p) + \sqrt n \ve (\hat \Lambda - \Lambda) + o_P(1)
\end{equation}
\end{Theorem}

The first matrix picks only off-diagonal elements of the left-hand side and the second one only diagonal elements. We shall now use this as well as the form of the asymptotic covariance matrix of the vectorized $\hat {\tilde \Sigma}$, i.e. $V_W$ to obtain limiting variance and covariances of eigenvalues and eigenvectors.

Due to the decomposition \eqref{equation:decompEq} we have, for $\BF_\Lambda$, the following relation between any off-diagonal element of $\hat{\tilde \Lambda}$ and the corresponding element in the estimate of eigenvectors, say $\hat {\tilde \Gamma}_\Lambda$:

$$ \sqrt n \hat {\tilde \gamma}_{\Lambda, ij} = \sqrt n \frac{\hat {\tilde \Lambda} (i,j) }{\tilde \lambda_i - \tilde \lambda_j}; \quad i \neq j$$

So that for eigenvector estimates of the original $\BF$ we have
%
\begin{equation}\label{equation:app1}
\sqrt n ( \hat{\tilde \gamma}_{i} - \gamma_i) =
\sqrt n \Gamma ( \hat {\tilde \gamma}_{\Lambda, i} - e_i ) =
\sqrt n \left[ \sum_{k=1; k \neq i}^p \hat {\tilde \gamma}_{\Lambda,i,k} \gamma_k + (\hat {\tilde \gamma}_{\Lambda,i,i} - 1) \gamma_i \right]
\end{equation}
%
Now $\sqrt n (\hat {\tilde \gamma}_{\Lambda,i,i} - 1) - 1) =  o_P(1)$ and $A\BV(\sqrt n \hat {\tilde \Lambda} (i,k), \sqrt n \hat {\tilde \Lambda} (i,l) ) = 0$ for $k \neq l$, so the above equation implies
%
$$ A\BV(g_i) = A \BV (\sqrt n ( \hat{\tilde \gamma}_{i} - \gamma_i)) = \sum_{k=1; k \neq i}^p \frac{A\BV(\sqrt n \hat {\tilde \Lambda} (i,k))}{(\tilde \lambda_i - \tilde \lambda_k)^2} \Gamma_k \Gamma_k^T $$
%

For the covariance terms, from \eqref{equation:app1} we get, for $i \neq j$,
%
\begin{eqnarray*}
A\BV(g_i, g_j) &=&
A\BV(\sqrt n ( \hat{\tilde \gamma}_{i} - \gamma_i),
\sqrt n ( \hat{\tilde \gamma}_{j} - \gamma_j))\\
&=& A\BV \left(
\sum_{k=1; k \neq i}^p \sqrt n \hat {\tilde \gamma}_{\Lambda,ik} \gamma_k,
\sum_{k=1; k \neq j}^p \sqrt n \hat {\tilde \gamma}_{\Lambda,jk} \gamma_k \right)\\
&=& A\BV \left(
\sqrt n \hat {\tilde \gamma}_{\Lambda,ik} \gamma_j,
\sqrt n \hat {\tilde \gamma}_{\Lambda,ik} \gamma_i \right)\\
&=& - \frac{A\BV(\sqrt n \hat {\tilde \Lambda} (i,j))}
{(\tilde \lambda_i - \tilde \lambda_j)^2} \gamma_j \gamma_i^T
\end{eqnarray*}
%
The exact forms given in the statement of the corollary now follows from the  Form of $V_W$ in Section~\ref{sec:appA}.

\paragraph{}For the on-diagonal elements of $\hat {\tilde \Lambda}$, using Theorem \ref{Theorem:decomp} we have for the $i^\text{th}$ eigenvalue of $\hat {\tilde \Lambda}$, say $\lambda_{\Lambda,i}$,
%
$$ \sqrt n \hat {\tilde \lambda}_{\Lambda,i} = \sqrt n \hat {\tilde \Lambda} (i,i), $$
%
for $i = 1,...,p$. Hence

\begin{eqnarray*}
A\BV(l_i) &=& A\BV( \sqrt n (\hat {\tilde \lambda}_{\Lambda,i} - \tilde \lambda_i) )\\
&=& A\BV( \sqrt n (\hat {\tilde \lambda}_{\Lambda,i} - \tilde \lambda_{\Lambda,i}) )\\
&=& A\BV(\sqrt n \hat {\tilde \Lambda} (i,i))
\end{eqnarray*}

A similar derivation gives the expression for $A\BV(l_i,l_j); i \neq j$. Finally, since the asymptotic covariance between an on-diagonal and an off-diagonal element of $\hat {\tilde \Lambda}$, it follows that the elements of $G_n$ and diagonal elements of $L_n$ are independent.
\end{proof}

\bibliographystyle{apalike}
\bibliography{WeightedSignBib_01_29_19}

\end{document}