### 2020-09-21 EJSSubmission/arxiv/arxiv-supp.tex

\appendix
\section*{Supplementary material}
\renewcommand{\thesection}{\Alph{section}}
\section{Form of $V_W$}\label{sec:appA}
First observe that for $F$ having covariance matrix $\Sigma = \Gamma\Lambda\Gamma^T$,
%
\begin{align}\label{eqn:VWeqn}
V_W  = (\Gamma \otimes \Gamma) V_{W, \Lambda} (\Gamma \otimes \Gamma)^T
\end{align}
%
where $V_{W, \Lambda}$ is the covariance matrix of $\BF_\Lambda$, the elliptic distribution with mean $\mu$ and covariance matrix $\Lambda$. Now,
%
\begin{eqnarray*}
V_{W, \Lambda} &=& \BE \left[ \ve \left\{ \frac{W^2 (Z, \BF_Z) \Lambda^{1/2} ZZ^T \Lambda^{1/2} }{ Z^T \Lambda Z} - \tilde \Lambda \right\} {\ve}^T \left\{ \frac{ W^2(Z, \BF_Z) \Lambda^{1/2} ZZ^T \Lambda^{1/2} }{ Z^T \Lambda Z} - \tilde \Lambda \right\} \right]\\
&=& \BE \left[ \ve \left\{ W^2 (Z, \BF_Z) \BS( \Lambda^{1/2} z; {\bf 0} ) \right\} {\ve}^T \left\{ W^2(Z, \BF_Z) \BS( \Lambda^{1/2}Z; \bf0) \right\} \right]\\
&& - \quad \ve( \tilde \Lambda) {\ve}^T( \tilde \Lambda )
\end{eqnarray*}

The matrix $\ve( \tilde\Lambda) \ve^T(\tilde\Lambda)$ consists of elements $\lambda_i\lambda_j$ at $(i,j)^\text{th}$ position of the $(i,j)^\text{th}$ block, and 0 otherwise. These positions correspond to variance and covariance components of on-diagonal elements. For the expectation matrix, all its elements are of the form $ \BE[ \sqrt{\lambda_a \lambda_b \lambda_c \lambda_d} Z_a Z_b Z_c Z_d . W^4 (Z, \BF_Z) / (Z^T \Lambda Z)^2]$, with $1 \leq a,b,c,d \leq p$. Since $W^4 (Z, \BF_Z) / (Z^T \Lambda Z)^2$ is even in $Z$, which has a spherically symmetric distribution, all such expectations will be 0 unless $a,b,c,d$ are all equal or pairwise equal. Following a similar derivation for spatial sign covariance matrices in \cite{ref:Biometrika14673_MagyarTyler}, we collect the non-zero elements and write the matrix of expectations:
%
$$ (\BI_{p^2} + \BK_{p,p}) \left\{ \sum_{a=1}^p \sum_{b=1}^p \tilde\gamma_{ab} (e_a e_a^T \otimes  e_b e_b^T) - \sum_{a=1}^p \tilde\gamma_{aa} (e_a e_a^T \otimes  e_a e_a^T) \right\} + \sum_{a=1}^p \sum_{b=1}^p \tilde\gamma_{ab} (e_a e_b^T \otimes  e_a e_b^T) $$
%
where $\BI_k = (e_1,...,e_k), \BK_{m,n} = \sum_{i=1}^m \sum_{j=1}^n \BJ_{ij} \otimes \BJ_{ij}^T$ with $\BJ_{ij} \in \BR^{m \times n}$ having 1 as $(i,j)$-th element and 0 elsewhere, and $\tilde\gamma_{mn} = \BE[ \lambda_m \lambda_n Z_m^2 Z_n ^2 . W^4 (Z, \BF_Z) / (Z^T \Lambda Z)^2]; 1 \leq m,n \leq p$.

\paragraph{}Putting everything together, denote by $\hat {\tilde \Lambda}$ the sample version of $\tilde \Lambda$, the weighted covariance matrix obtained from $\BF_\Lambda$, i.e. $\hat {\tilde \Lambda} = \sum_{i=1}^n W_n^2 (Z_i, \BF_{Z,n}) \BS( \Lambda^{1/2} Z_i; \hat \mu_n)/n $. Then the different types of elements in the matrix $\hat {\tilde \Lambda}$ are as given below ($1 \leq a,b,c,d \leq p$):

\begin{itemize}
\item Variance of on-diagonal elements
%
$$ A\BV( \sqrt n \hat {\tilde \Lambda} (a,a) ) = \BE \left[ \frac{ W^4(Z, \BF_Z) \lambda_a^2 Z_a^4}{(Z^T \Lambda Z)^2} \right] - \tilde\lambda_a^2 $$

\item Variance of off-diagonal elements ($a \neq b$)
%
$$ A\BV( \sqrt n \hat {\tilde \Lambda} (a,b) ) = \BE \left[ \frac{W^4 (Z, \BF_Z) \lambda_a \lambda_b Z_a^2 Z_b^2}{(Z^T \Lambda Z)^2} \right] $$

\item Covariance of two on-diagonal elements ($a \neq b$)
%
$$ A\BV(\sqrt n \hat {\tilde \Lambda} (a,a), \sqrt n \hat {\tilde \Lambda} (b,b) )
= \BE \left[ \frac{W^4 (Z, \BF_Z) \lambda_a \lambda_b Z_a^2 Z_b^2}{(Z^T \Lambda Z)^2} \right] - \tilde\lambda_{a} \tilde\lambda_{b} $$

\item Covariance of two off-diagonal elements ($a \neq b, c \neq d$)
%
$$ A\BV(\sqrt n \hat {\tilde \Lambda} (a,b), \sqrt n \hat {\tilde \Lambda} (c,d) ) = 0 $$

\item Covariance of one off-diagonal and one on-diagonal element ($a \neq b \neq c$)
%
$$ A\BV(\sqrt n \hat {\tilde \Lambda} (a,b), \sqrt n \hat {\tilde \Lambda} (c,c) ) = 0 $$
\end{itemize}

\noindent The above give all the elements of $V_{W,\Lambda}$. We plug these in \eqref{eqn:VWeqn} to recover $V_W$.

\section{Proofs}\label{section:appB}

\begin{proof}[Proof sketch of Corollary 2.4]
%A first step to obtain asymptotic normality for the high-dimensional location test statistic $C_{n,w}$ is obtaining an equivalent result of Lemma 2.1 in \cite{ref:JASA151658_WangPengLi}:
%
%\begin{Lemma}\label{Lemma:HDlemma21} Consider a random variable $\epsilon$ having an elliptical distribution with mean ${\bf 0}$ and covariance matrix $\Lambda$. Then, under the conditions 1-2 in Corollary 2.4, we have
%%
%\begin{eqnarray}
%\BE[ (\epsilon_{W1}^T \epsilon_{W2})^4 ] &=& O(1) \BE^2[ (\epsilon_{W1}^T \epsilon_{W2})^2 ]\\
%\BE[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W1})^2 ] &=& O(1) \BE^2[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W1})^2 ]\\
%\BE[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W2})^2 ] &=& o(1) \BE^2[ (\epsilon_{W1}^T \Psi_{1W} \epsilon_{W1})^2 ]
%\end{eqnarray}
%%
%with $\epsilon_W = W(\epsilon, \BF_\epsilon) S(\epsilon)$ where $\BF_\epsilon$ is the distribution of $\epsilon$.
%\end{Lemma}
%%
%A proof of this lemma is derived using results in section 3 of \cite{ElKaroui09}, noticing that any-scalar valued 1-Lipschitz function of $\epsilon_W$ is a $W_\max$-Lipschitz function of $S(\epsilon)$.

%To derive the asymptotic distribution under contiguous alternatives we need the conditions (C3)-(C6) in \cite{ref:JASA151658_WangPengLi}, as well as
We start with slightly modified versions of Lemmas A.4 and A.5 in \cite{ref:JASA151658_WangPengLi}:

\begin{Lemma}\label{lemma:lemmaB1}
Given that condition 1 in Corollary 2.4 holds, we have $\lambda_{\max} (\Psi_{1W}) \leq 2 W_{\max} \frac{\lambda_{\max} (\Sigma) }{\text{Tr} (\Sigma)} (1+o(1))$.
\end{Lemma}

\begin{Lemma}\label{lemma:lemmaB2}
Define $\Psi_{3W} = \BE \left[ \frac{W^2(\epsilon, \BF_\epsilon)}{| \epsilon |^2} (\BI_p - S(\epsilon) S(\epsilon)^T )\right] $. Then $\lambda_{\max} (\Psi_{2W}) \leq W_{\max} \BE(| \epsilon |^{-1})$ and $\lambda_{\max} (\Psi_{3W}) \leq (W_{\max})^2 \BE( | \epsilon|^{-2})$. Further, if conditions 1 and 2 of Corollary 2.4 hold then $\lambda_{\min} (\Psi_{2W}) \geq \BE( W(\epsilon, \BF_\epsilon)/ | \epsilon |)(1+o(1))/\sqrt 3$.
\end{Lemma}
%
The lemmas can be proved using similar steps as the proofs of the original lemmas in \cite{ref:JASA151658_WangPengLi} and using the upper bound on the weight function. Corollary 2.4 is now proved by applying Corollary 2.2, plugging in upper bound of $\lambda_{\max} (\Psi_{1W})$ from Lemma~\ref{lemma:lemmaB1} and lower bound of $\lambda_{\max} (\Psi_{1W})$ from Lemma~\ref{lemma:lemmaB2} into the ARE expression.
\end{proof}

\begin{proof}[Proof of Theorem 3.4]
We suppose $G_n = (g_1, \ldots, g_p), L_n = \diag(l_1, \ldots, l_p)$. In spirit, this corollary is similar to Theorem 13.5.1 in \cite{ref:AndersonBook09}. We start with the following result, due to \citep{ref:SPL12765_Taskinenetal}, allows us to obtain asymptotic joint distributions of eigenvectors and eigenvalues of $\hat{\tilde \Sigma}$, provided we know the limiting distribution of $\hat{\tilde \Sigma}$ itself:

\begin{Theorem} \label{Theorem:decomp} 
Let $\BF_\Lambda$ be defined as before, and $\hat C$ be any positive definite symmetric $p \times p$ matrix such that at $F_\Lambda$ the limiting distribution of $\sqrt n \ve(\hat C - \Lambda)$ is a $p^2$-variate (singular) normal distribution with mean zero. Write the spectral decomposition of $\hat C$ as $\hat C = \hat P \hat\Lambda \hat P^T$. Then the limiting distributions of $\sqrt n \ve(\hat P - \BI_p)$ and $\sqrt n \ve( \hat\Lambda - \Lambda)$ are multivariate (singular) normal and
%
\begin{equation} \label{equation:decompEq}
\sqrt n \ve (\hat C - \Lambda)  = \left[ (\Lambda \otimes \BI_p) - (\BI_p \otimes \Lambda) \right] \sqrt n \ve (\hat P - \BI_p) + \sqrt n \ve (\hat \Lambda - \Lambda) + o_P(1)
\end{equation}
\end{Theorem}

The first matrix picks only off-diagonal elements of the left-hand side and the second one only diagonal elements. We shall now use this as well as the form of the asymptotic covariance matrix of the vectorized $\hat {\tilde \Sigma}$, i.e. $V_W$ to obtain limiting variance and covariances of eigenvalues and eigenvectors.

Due to the decomposition \eqref{equation:decompEq} we have, for $\BF_\Lambda$, the following relation between any off-diagonal element of $\hat{\tilde \Lambda}$ and the corresponding element in the estimate of eigenvectors, say $\hat {\tilde \Gamma}_\Lambda$:

$$ \sqrt n \hat {\tilde \gamma}_{\Lambda, ij} = \sqrt n \frac{\hat {\tilde \Lambda} (i,j) }{\tilde \lambda_i - \tilde \lambda_j}; \quad i \neq j$$

So that for eigenvector estimates of the original $\BF$ we have
%
\begin{equation}\label{equation:app1}
\sqrt n ( \hat{\tilde \gamma}_{i} - \gamma_i) =
\sqrt n \Gamma ( \hat {\tilde \gamma}_{\Lambda, i} - e_i ) =
\sqrt n \left[ \sum_{k=1; k \neq i}^p \hat {\tilde \gamma}_{\Lambda,i,k} \gamma_k + (\hat {\tilde \gamma}_{\Lambda,i,i} - 1) \gamma_i \right]
\end{equation}
%
Now $\sqrt n (\hat {\tilde \gamma}_{\Lambda,i,i} - 1) =  o_P(1)$ and $A\BV(\sqrt n \hat {\tilde \Lambda} (i,k), \sqrt n \hat {\tilde \Lambda} (i,l) ) = 0$ for $k \neq l$, so the above equation implies
%
$$ A\BV(g_i) = A \BV (\sqrt n ( \hat{\tilde \gamma}_{i} - \gamma_i)) = \sum_{k=1; k \neq i}^p \frac{A\BV(\sqrt n \hat {\tilde \Lambda} (i,k))}{(\tilde \lambda_i - \tilde \lambda_k)^2} \gamma_k \gamma_k^T $$
%

For the covariance terms, from \eqref{equation:app1} we get, for $i \neq j$,
%
\begin{eqnarray*}
A\BV(g_i, g_j) &=&
A\BV(\sqrt n ( \hat{\tilde \gamma}_{i} - \gamma_i),
\sqrt n ( \hat{\tilde \gamma}_{j} - \gamma_j))\\
&=& A\BV \left(
\sum_{k=1; k \neq i}^p \sqrt n \hat {\tilde \gamma}_{\Lambda,ik} \gamma_k,
\sum_{k=1; k \neq j}^p \sqrt n \hat {\tilde \gamma}_{\Lambda,jk} \gamma_k \right)\\
&=& A\BV \left(
\sqrt n \hat {\tilde \gamma}_{\Lambda,ik} \gamma_j,
\sqrt n \hat {\tilde \gamma}_{\Lambda,ik} \gamma_i \right)\\
&=& - \frac{A\BV(\sqrt n \hat {\tilde \Lambda} (i,j))}
{(\tilde \lambda_i - \tilde \lambda_j)^2} \gamma_j \gamma_i^T
\end{eqnarray*}
%
The exact forms given in the statement of the corollary now follows from the  Form of $V_W$ in Section~\ref{sec:appA}.

\paragraph{}For the on-diagonal elements of $\hat {\tilde \Lambda}$, using Theorem \ref{Theorem:decomp} we have for the $i^\text{th}$ eigenvalue of $\hat {\tilde \Lambda}$, say $\lambda_{\Lambda,i}$,
%
$$ \sqrt n \hat {\tilde \lambda}_{\Lambda,i} = \sqrt n \hat {\tilde \Lambda} (i,i), $$
%
for $i = 1,...,p$. Hence

\begin{eqnarray*}
A\BV(l_i) &=& A\BV( \sqrt n (\hat {\tilde \lambda}_{\Lambda,i} - \tilde \lambda_i) )\\
&=& A\BV( \sqrt n (\hat {\tilde \lambda}_{\Lambda,i} - \tilde \lambda_{\Lambda,i}) )\\
&=& A\BV(\sqrt n \hat {\tilde \Lambda} (i,i))
\end{eqnarray*}

A similar derivation gives the expression for $A\BV(l_i,l_j); i \neq j$. Finally, since the asymptotic covariance between an on-diagonal and an off-diagonal element of $\hat {\tilde \Lambda}$, it follows that the elements of $G_n$ and diagonal elements of $L_n$ are independent.
\end{proof}